### 2015-10-28 proposal.tex
Project Proposal

\documentclass{article}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{courier}
\usepackage{multicol}
\usepackage{microtype} %required to fix courier linebreaking issues
\usepackage{hyperref} % make references into links
\usepackage{amsfonts}
\hypersetup{colorlinks=true, linkcolor=blue} %make them pretty links
\usepackage[all]{hypcap} %make them link to the right place

\begin{document}
	\title{CS182: NBA Scheduler}
	\author{Abhishek Malali, Charles Liu (cliu02@g.harvard.edu), Samuel Dalton}
	\date{\today}
	\maketitle

	\section{Introduction}

	The National Basketball Association is a 30 team league that plays an 82 game schedule over a 6-month period. It showcases some of the best athletes in the world and has generated globally recognized brands. As such, it is a multi-billion dollar industry and is constantly looking for ways to improve. One such way recently has been to reorganize it's schedule.

	\subsection{NBA Scheduling Formula}

	Currently there are two conferences (East, West) with three divisions of five teams within each conference. The 82-game schedule is currently set up as follows:

	\begin{enumerate}
		\item 4 games against the other 4 division opponents, [4x4=16 games]
		\item 4 games against 6 (out-of-division) conference opponents, [4x6=24 games]
		\item 3 games against the remaining 4 conference teams, [3x4=12 games]
		\item 2 games against teams in the opposing conference. [2x15=30 games]
	\end{enumerate}

	And of course, each team must play 42 home and away games.

	\subsection{Court Availability}

	As these are actual games with required venues, each team must provide the following:

	\begin{enumerate}
		\item At least 50 dates on which their home court will be available
		\item 4 Mondays
		\item 4 Thursdays (to help TNT plan its telecasts).
	\end{enumerate}

	\subsection{Official Breaks}

	There are no games on the following days:

	\begin{enumerate}
		\item Christmas Eve
		\item All-Star Week
		\item NCAA Men's Division I Basketball Championship Game
	\end{enumerate}

	\section{Project Goals}

	While the NBA has obviously been able to put out a schedule every year for every team, there has been recent criticism - mainly that there are too many ``back-to-back'' stretches, where teams potentially play 4 games in 5 nights, and the length of travel in a short amount of time is too taxing on the players. Our end-goal would be to address these concerns.

	\subsection{Data Generation}
	For the data - we know what the 30 teams are and will calculate the exact distance between any two team venues. Then, for a specific year we will randomly generate the court availability constraints mentioned above as well as the week of the All-Star break (this must be a week in February) and NCAA Championship Game (First Monday of April).

	\subsection{Initial Schedule}
	We will then model this problem as a CSP similar to ones we've seen in class and generate an initial solution. We will also implement some local search algorithms and test how feasible they are given the number of constraints.

	\subsection{More Player Friendly Schedule}
	We will then look to address the player concerns that there are too many games in too short a time. We'll have generic variables $N$ and $M$ and add the following constraint that no team can play $N$ games in $M$ nights.  Additionally, we want to find the solution that would minimize the travel distance of a team - this lessens the wear on players and saves flight coordination issues. This could potentially be an increasingly computationally intensive problem - so we will look to see if certain optimizations can provide a solution:

	\begin{enumerate}
		\item Some form of greedy next-best search in choosing the next game - choose the next game for a team that makes sure this constraint is met then perform arc consistency
		\item Generate some heuristic based on the number of games left against divisional opponents (who would be geographically closer) to perform some kind of A* search
		\item Research online for other techniques that haven't been discussed in class - surprisingly this has been an issue well researched in the NFL but not the NBA (probably because that's a smaller problem with only 16 games), but perhaps we can gain some insights from reading papers optimizing their schedule
	\end{enumerate}

	\section{Group Member Roles}
	The project steps are closely related, so each member's roles will certainly overlap. However, a list of requirements would be:

	\begin{enumerate}
		\item Data generation
		\item Initial model formulation
		\item Initial CSP search code
		\item Local Search code
		\item Heuristic Function on travel
		\item Model revisited including new constraints
		\item Find relevant research papers for NFL
	\end{enumerate}

\end{document}

### 2015-10-28 proposal.tex
updates

@@
 		\item Some form of greedy next-best search in choosing the next game - choose the next game for a team that makes sure this constraint is met then perform arc consistency
-		\item Generate some heuristic based on the number of games left against divisional opponents (who would be geographically closer) to perform some kind of A* search
+		\item The minimizing distance problem is classic TSP - could try to model this problem in form of A* and find a heuristic based on number of divisional games remaining (divisional games are regionally closer)
 		\item Research online for other techniques that haven't been discussed in class - surprisingly this has been an issue well researched in the NFL but not the NBA (probably because that's a smaller problem with only 16 games), but perhaps we can gain some insights from reading papers optimizing their schedule
@@
 
+	\section{Resources}
+	\begin{itemize}
+		\item \url{http://www.public.asu.edu/~huanliu/AI04S/project1.htm}
+		\item \url{http://www.cc.gatech.edu/~bdilkina/papers/NFLsched1.pdf}
+		\item \url{http://www.sloansportsconference.com/wp-content/uploads/2015/02/SSAC15-RP-Finalist-Alleviating-competitive-imbalance-in-NFL-schedules.pdf}
+	\end{itemize}
+
 \end{document}

### 2015-10-28 proposal.tex
additional clarifications on algorithms

@@
 	We will then look to address the player concerns that there are too many games in too short a time.
-We'll have generic variables $N$ and $M$ and add the following constraint that no team can play $N$ games in $M$ nights.
-Additionally, we want to find the solution that would minimize the travel distance of a team - this lessens the wear on players and saves flight coordination issues.
+We'll have generic variables $N$ and $M$ and add the following constraint that a team can play at most $N$ games in $M$ nights.
+If time permits, we want to find the solution that would minimize the total distance traveled - this lessens the wear on players and saves flight coordination issues.
 This could potentially be an increasingly computationally intensive problem - so we will look to see if certain optimizations can provide a solution:
@@
 	\begin{enumerate}
-		\item Some form of greedy next-best search in choosing the next game - choose the next game for a team that makes sure this constraint is met then perform arc consistency
-		\item The minimizing distance problem is classic TSP - could try to model this problem in form of A* and find a heuristic based on number of divisional games remaining (divisional games are regionally closer)
+		\item Some form of greedy next-best search in choosing the next game - choose the next game for a team that makes sure the $N$ games in $M$ nights constraint is met then perform arc consistency
+		\item The minimizing distance problem for a specific team is classic TSP - could try to model this problem in form of A* and find a heuristic based on number of divisional games remaining (divisional games are regionally closer) - then rearrange schedule to fit that team's path.
+We could try to first accomplish this but the goal is to minimize the combined distance for every team traveled which will require outside research to see if this is even feasible
 		\item Research online for other techniques that haven't been discussed in class - surprisingly this has been an issue well researched in the NFL but not the NBA (probably because that's a smaller problem with only 16 games), but perhaps we can gain some insights from reading papers optimizing their schedule

### 2015-10-28 proposal.tex
fix typo
Mostly grammatical changes

@@
 	\title{CS182: NBA Scheduler}
-	\author{Abhishek Malali, Charles Liu (cliu02@g.harvard.edu), Samuel Dalton}
+	\author{Abhishek Malali, Charles Liu (cliu02@g.harvard.edu), Samuel Daulton}
 	\date{\today}
@@
 
-	The National Basketball Association is a 30 team league that plays an 82 game schedule over a 6-month period.
+	The National Basketball Association is a 30 team league that organizes an 82 game schedule for each team over a 6-month period.
 It showcases some of the best athletes in the world and has generated globally recognized brands.
 As such, it is a multi-billion dollar industry and is constantly looking for ways to improve.
-One such way recently has been to reorganize it's schedule.
+Recently, one such way has been to reorganize its schedule.
 
@@
 
-	Currently there are two conferences (East, West) with three divisions of five teams within each conference. The 82-game schedule is currently set up as follows:
+	Currently there are two conferences (East, West) each comprised of three divisions each of which contain five teams. The 82-game schedule is currently set up as follows:
 
@@
 
-	And of course, each team must play 42 home and away games.
+	And of course, each team must play 41 home and 41 away games.
 
@@
 
+    \subsection{Additional Assumptions}
+
+    There are no non-NBA events in the venues that may further constrain a team's schedule.
+
 	\section{Project Goals}
 
-	While the NBA has obviously been able to put out a schedule every year for every team, there has been recent criticism - mainly that there are too many ``back-to-back'' stretches, where teams potentially play 4 games in 5 nights, and the length of travel in a short amount of time is too taxing on the players.
-Our end-goal would be to address these concerns.
+	While the NBA has obviously been able to put out a schedule every year for every team, there has been recent criticism; the main criticism is there are too many ``back-to-back'' stretches, where teams potentially play 4 games in 5 nights.
+Moreover, the length of travel in a short amount of time is too taxing on the players.
+Our end-goal would be to minimize these schedule-induced "costs" on a player's energy level.
 
@@
 	\subsection{Initial Schedule}
-	We will then model this problem as a CSP similar to ones we've seen in class and generate an initial solution.
+	We will then model this problem as a CSP, similar to ones we've seen in class, and generate an initial solution.
 We will also implement some local search algorithms and test how feasible they are given the number of constraints.
@@
 	We will then look to address the player concerns that there are too many games in too short a time.
-We'll have generic variables $N$ and $M$ and add the following constraint that a team can play at most $N$ games in $M$ nights.
+We'll add a new constraint to model this concern: using generic variables $N$ and $M$, we would add the constraint that a team can play at most $N$ games in $M$ nights.
 If time permits, we want to find the solution that would minimize the total distance traveled - this lessens the wear on players and saves flight coordination issues.
@@
 		\item Some form of greedy next-best search in choosing the next game - choose the next game for a team that makes sure the $N$ games in $M$ nights constraint is met then perform arc consistency
-		\item The minimizing distance problem for a specific team is classic TSP - could try to model this problem in form of A* and find a heuristic based on number of divisional games remaining (divisional games are regionally closer) - then rearrange schedule to fit that team's path.
+		\item The minimizing distance problem for a specific team is classic CSP - could try to model this problem in form of A* and find a heuristic based on number of divisional games remaining (divisional games are regionally closer) - then rearrange schedule to fit that team's path.
 We could try to first accomplish this but the goal is to minimize the combined distance for every team traveled which will require outside research to see if this is even feasible

### 2015-10-28 proposal.tex
change tsp to traveling salesman problem explicitly

@@
 		\item Some form of greedy next-best search in choosing the next game - choose the next game for a team that makes sure the $N$ games in $M$ nights constraint is met then perform arc consistency
-		\item The minimizing distance problem for a specific team is classic CSP - could try to model this problem in form of A* and find a heuristic based on number of divisional games remaining (divisional games are regionally closer) - then rearrange schedule to fit that team's path.
+		\item The minimizing distance problem for a specific team is classic traveling salesman problem - could try to model this problem in form of A* and find a heuristic based on number of divisional games remaining (divisional games are regionally closer) - then rearrange schedule to fit that team's path.
 We could try to first accomplish this but the goal is to minimize the combined distance for every team traveled which will require outside research to see if this is even feasible

### 2015-10-28 proposal.tex
update goal

@@
 		\item The minimizing distance problem for a specific team is classic traveling salesman problem - could try to model this problem in form of A* and find a heuristic based on number of divisional games remaining (divisional games are regionally closer) - then rearrange schedule to fit that team's path.
-We could try to first accomplish this but the goal is to minimize the combined distance for every team traveled which will require outside research to see if this is even feasible
+We could try to first accomplish this but the goal is to minimize the combined distance for every team traveled which will require outside research to see if this is even feasible.
+At a minimum we will try to beat the total distance traveled from the actual NBA schedule for this year.
 		\item Research online for other techniques that haven't been discussed in class - surprisingly this has been an issue well researched in the NFL but not the NBA (probably because that's a smaller problem with only 16 games), but perhaps we can gain some insights from reading papers optimizing their schedule

### 2015-10-29 proposal.tex
update layout

@@
 \usepackage[all]{hypcap} %make them link to the right place
+\usepackage{url}
+\makeatletter
+\g@addto@macro{\UrlBreaks}{\UrlOrds}
+\makeatother
+\usepackage{geometry}
+\geometry{legalpaper, portrait, margin=1in}
 
@@
 	\title{CS182: NBA Scheduler}
-	\author{Abhishek Malali, Charles Liu (cliu02@g.harvard.edu), Samuel Daulton}
+	\author{Abhishek Malali (abhishekmalali@g.harvard.edu)\\
+			Charles Liu (cliu02@g.harvard.edu)\\
+			Samuel Daulton (sdaulton@g.harvard.edu)}
 	\date{\today}
@@
 		\item Some form of greedy next-best search in choosing the next game - choose the next game for a team that makes sure the $N$ games in $M$ nights constraint is met then perform arc consistency
-		\item The minimizing distance problem for a specific team is classic traveling salesman problem - could try to model this problem in form of A* and find a heuristic based on number of divisional games remaining (divisional games are regionally closer) - then rearrange schedule to fit that team's path.
+		\item The minimizing distance problem for a specific team is similar to traveling salesman problem - could try to model this problem in form of A* and find a heuristic based on number of divisional games remaining (divisional games are regionally closer) - then rearrange schedule to fit that team's path.
 We could try to first accomplish this but the goal is to minimize the combined distance for every team traveled which will require outside research to see if this is even feasible.
-At a minimum we will try to beat the total distance traveled from the actual NBA schedule for this year.
+At a minimum we will try to beat the total distance traveled from the actual NBA schedule for this year through a greedy approach of finding minimum distance traveled while optimizing for 1 team, then minimum distance traveled for for 2 teams including the best from the 1-team, and so on until as many teams as possible have minimized their travel distance.
 		\item Research online for other techniques that haven't been discussed in class - surprisingly this has been an issue well researched in the NFL but not the NBA (probably because that's a smaller problem with only 16 games), but perhaps we can gain some insights from reading papers optimizing their schedule

### 2015-10-29 proposal.tex
fix typo

--- 2015-10-28 proposal.tex
@@
 		\item Some form of greedy next-best search in choosing the next game - choose the next game for a team that makes sure the $N$ games in $M$ nights constraint is met then perform arc consistency
-		\item The minimizing distance problem for a specific team is classic traveling salesman problem - could try to model this problem in form of A* and find a heuristic based on number of divisional games remaining (divisional games are regionally closer) - then rearrange schedule to fit that team's path.
-We could try to first accomplish this but the goal is to minimize the combined distance for every team traveled which will require outside research to see if this is even feasible
+		\item The minimizing distance problem for a specific team is similar to traveling salesman problem - try to model this problem in form of A* and find a heuristic based on number of divisional games remaining (divisional games are regionally closer) - then rearrange schedule to fit that team's path.
+The goal is to minimize the combined distance for every team traveled which will require outside research to see if this is even feasible.
+At a minimum we will try to beat the total distance traveled from the actual NBA schedule for this year through a greedy approach of finding minimum distance traveled while optimizing for 1 team, then minimum distance traveled for 2 teams starting with the best from the 1-team, and so on until as many teams as possible have minimized their travel distance.
 		\item Research online for other techniques that haven't been discussed in class - surprisingly this has been an issue well researched in the NFL but not the NBA (probably because that's a smaller problem with only 16 games), but perhaps we can gain some insights from reading papers optimizing their schedule

### 2015-12-05 constraint_changed_venues.py
Changed valid_venue constraint, added binary search for dates csp, created successor_domains method for venues, changed next game selection when assigning venues for be the one with minimum game number.  Note: I created new files for the time being with my changes to constraint and scheduler

@@ -1,0 +1,193 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sat Nov 21 17:23:47 2015
+
+Check factors for:
+1) For each game, each team/opponent home/away lines up
+2) Games against divisional opponents not exceeded, home/away even
+3) Games against inter-conference opponents not exceeded, home/away even
+4) Games against conf_opponents not exceeded, home/away even
+5) Games against conference opponents not in conf_opponents not exceeded, home/away as even as possible
+
+@author: charlesliu
+"""
+
+def add(m, k):
+    if k in m:
+        m[k] += 1
+    else:
+        m[k] = 1
+
+#total games two teams should play, pass in m which is a TGBase object or child
+def total_games(m, team, opponent):
+    if opponent.division is team.division:
+        return m.DIVISIONAL_GAMES
+    elif opponent in team.conf_opponents:
+        return m.CONF_OPPONENTS_GAMES
+    elif opponent.division.conference is team.division.conference:
+        return m.REMAINING_CONF_GAMES
+    else:
+        return m.INTERCONF_GAMES
+        
+def valid_date(m, sk):
+    domains = m.states[m.domains]
+    selected = m.states[m.selected]
+    t1, t2, g_n = sk
+    dateindex = selected[sk]
+    #constraints are:
+    #any game before g_n can't have a date >= g_n
+    #any game after g_n can't have a date <= g_n
+    
+    for state in domains:
+        x1, x2, g_x = state
+        #involves one of the two teams playing but not sk
+        if state is not sk and (t1 in [x1, x2] or t2 in [x1, x2]):
+            if g_x < g_n:
+                #remove all dates >= dateindex
+                domains[state][:] = [x for x in domains[state] if x < dateindex]
+            elif g_x > g_n:
+                #remove all dates <= dateindex
+                domains[state][:] = [x for x in domains[state] if x > dateindex]
+            if len(domains[state]) == 0 and selected[state] is None:
+                return False
+        elif state is sk:
+            domains[state][:] = []
+    
+    return True
+        
+#check that the total number of home games is TOTAL_GAMES/2
+def valid_venue(m, sk):
+    selected = m.states[m.selected]
+    domains = m.states[m.domains]
+    home_games = {}
+    away_games = {}
+    num_games = {}
+    # NOTE: we are only concerned with t1 and t2 since those are the only teams that are
+    # affected by the selected game.
+    t1, t2, gn = sk
[…]

### 2015-12-05 domain.py

@@ -65,7 +65,7 @@ class TGBase(object):
                     new_obj = self.copy_states(domain_class, dk, sk, dom_elem)
                     if constraint_func(new_obj, sk):
                         domains.append(new_obj)
-                    added_elems.add(dom_elem)
+                    added_elems.add(dom_elem) 
             return domains
         return None
     
@@ -140,11 +140,15 @@ class Venues(TGBase):
         return self.successorDomains(Venues, constraint.valid_venue)
     def min_key_helper(self, min_k):
         tx, ty = min_k
+        # Find matchup between tx, ty with the lowest game number
+        min_gn = float('inf')
+        min_state = None
         for state in self.states[self.selected]:
             t1, t2, gn = state
-            if self.states[self.selected][state] is None and t1 is tx and t2 is ty:
-                return state
-        return None
+            if self.states[self.selected][state] is None and t1 is tx and t2 is ty and gn < min_gn:
+                    min_gn = gn
+                    min_state = state      
+        return min_state
     def min_key(self):
         teams_selected = {}
         for state in self.states[self.selected]:
@@ -158,7 +162,7 @@ class Venues(TGBase):
                     teams_selected.pop(t2, None)
         
         #teams_selected stores how many teams have selected their home/away for games
-        #removing those that have had all their games scheduled with hom/away
+        #removing those that have had all their games scheduled with home/away
         #if it's empty then no team has had a selected home/away yet
         if len(teams_selected) == 0:
             for k in self.states[self.domains]:
@@ -178,7 +182,30 @@ class Venues(TGBase):
                 best_len = l
                 best_k = k
         return (best_k, self.min_key_helper(best_k)) if best_k is not None else (None, None)     
-        
+
+    def successorDomains(self, domain_class, constraint_func):
+        domains = []
+        dk, sk = self.min_key()
+        if dk is not None and sk is not None:
+            #iterate through domain and add to list but keep track of ones we've already done
+            #because some domains are repeated elements
+            # Note for venues, dom_elems are either True or False,
+            # so we can stop traversing self.ordered_domain(dk, sk) if we find both True and False
+            added_elems = set()
+            for dom_elem in self.ordered_domain(dk, sk):
+                if dom_elem not in added_elems:
+                    new_obj = self.copy_states(domain_class, dk, sk, dom_elem)
+                    # all this constraint does is remove domains
+                    if constraint_func(new_obj, sk):
+                        domains.append(new_obj)
+                    added_elems.add(dom_elem)
[…]

### 2015-12-05 constraint.py
Solved Venues CSP

@@ -62,6 +62,61 @@ def valid_venue(m, sk):
     home_games = {}
     away_games = {}
     num_games = {}
+    # NOTE: we are only concerned with t1 and t2 since those are the only teams that are
+    # affected by the selected game.
+    # Note, since we ordered when generating team pairs, t1 < t2. So if t1, t2 are both in a game,
+    # the order is (t1,t2), not (t2,t1)
+    t1, t2, gn = sk
+    for state in selected:
+        tx, ty, gnxy = state
+        if (t1 in [tx,ty] or t2 in [tx,ty]) and (selected[state] is not None):
+            home = selected[state]
+            
+            
+            if t1 == tx:
+                if t2 == ty:
+                    add(num_games, (t1, t2))
+                    add(away_games if home else home_games, t2)
+                add(home_games if home else away_games, t1)
+            elif t1 == ty:
+                add(away_games if home else home_games, t1)
+            
+            # t1 not in state
+            elif t2 == tx:
+                add(home_games if home else away_games, t2)
+            #t2 == ty
+            else:
+                add(away_games if home else home_games, t2)   
+      
+    # scheduled the required number of games between the two teams
+    # clear domain
+    if num_games[(t1,t2)] == total_games(m, t1, t2):
+        domains[(t1,t2)] = []
+        return True
+    
+    # add t1 and t2 to list if respectively if they have played enough home games or away games
+    enough_home_games_list = []
+    enough_away_games_list = []
+    for team in [t1,t2]:
+        if team in home_games and home_games[team] == (m.TOTAL_GAMES+1)/2:
+            enough_home_games_list.append(team)
+        if team in away_games and away_games[team] == (m.TOTAL_GAMES+1)/2:
+            enough_away_games_list.append(team)
+    # iterate over team pairs and remove home games for teams (t1 or t2) that have played enough
+    # home games or away games
+    for team_pair in domains:
+        tx, ty = team_pair
+        if tx in enough_home_games_list:
+            domains[team_pair] = [x for x in domains[team_pair] if not x]
+        elif ty in enough_home_games_list:
+            domains[team_pair] = [x for x in domains[team_pair] if x]
+        if tx in enough_away_games_list:
+            domains[team_pair] = [x for x in domains[team_pair] if x] 
+        elif ty in enough_away_games_list:
+            domains[team_pair] = [x for x in domains[team_pair] if not x]
+    return True
+    #old method below
+    '''
     for state in selected:
[…]

### 2015-12-05 constraint_changed_venues.py

@@ -1,193 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Sat Nov 21 17:23:47 2015
-
-Check factors for:
-1) For each game, each team/opponent home/away lines up
-2) Games against divisional opponents not exceeded, home/away even
-3) Games against inter-conference opponents not exceeded, home/away even
-4) Games against conf_opponents not exceeded, home/away even
-5) Games against conference opponents not in conf_opponents not exceeded, home/away as even as possible
-
-@author: charlesliu
-"""
-
-def add(m, k):
-    if k in m:
-        m[k] += 1
-    else:
-        m[k] = 1
-
-#total games two teams should play, pass in m which is a TGBase object or child
-def total_games(m, team, opponent):
-    if opponent.division is team.division:
-        return m.DIVISIONAL_GAMES
-    elif opponent in team.conf_opponents:
-        return m.CONF_OPPONENTS_GAMES
-    elif opponent.division.conference is team.division.conference:
-        return m.REMAINING_CONF_GAMES
-    else:
-        return m.INTERCONF_GAMES
-        
-def valid_date(m, sk):
-    domains = m.states[m.domains]
-    selected = m.states[m.selected]
-    t1, t2, g_n = sk
-    dateindex = selected[sk]
-    #constraints are:
-    #any game before g_n can't have a date >= g_n
-    #any game after g_n can't have a date <= g_n
-    
-    for state in domains:
-        x1, x2, g_x = state
-        #involves one of the two teams playing but not sk
-        if state is not sk and (t1 in [x1, x2] or t2 in [x1, x2]):
-            if g_x < g_n:
-                #remove all dates >= dateindex
-                domains[state][:] = [x for x in domains[state] if x < dateindex]
-            elif g_x > g_n:
-                #remove all dates <= dateindex
-                domains[state][:] = [x for x in domains[state] if x > dateindex]
-            if len(domains[state]) == 0 and selected[state] is None:
-                return False
-        elif state is sk:
-            domains[state][:] = []
-    
-    return True
-        
-#check that the total number of home games is TOTAL_GAMES/2
-def valid_venue(m, sk):
-    selected = m.states[m.selected]
-    domains = m.states[m.domains]
-    home_games = {}
-    away_games = {}
-    num_games = {}
-    # NOTE: we are only concerned with t1 and t2 since those are the only teams that are
-    # affected by the selected game.
-    t1, t2, gn = sk
[…]

### 2015-12-05 domain.py

@@ -149,6 +149,7 @@ class Venues(TGBase):
                     min_gn = gn
                     min_state = state      
         return min_state
+    
     def min_key(self):
         teams_selected = {}
         for state in self.states[self.selected]:
@@ -172,13 +173,13 @@ class Venues(TGBase):
         #choose the team with the most selected home/away venues, i.e. least domain size remaining
         min_k = sorted(teams_selected.keys(), key=lambda x: -teams_selected[x])[0]
         best_k = None
-        best_len = float("inf")
+        best_len = float("-inf")
         #we have the team we want to next choose a venue for, but domains are pairs of teams, 
-        #so choose the pair with the remaining least domain size
+        #so choose the pair with the largest remaining domain size
         for k in self.states[self.domains]:
             t1, t2 = k
             l = len(self.states[self.domains][k])
-            if (t1 is min_k or t2 is min_k) and  l > 0 and l < best_len:
+            if (t1 is min_k or t2 is min_k) and  l > 0 and l > best_len:
                 best_len = l
                 best_k = k
         return (best_k, self.min_key_helper(best_k)) if best_k is not None else (None, None)     

### 2015-12-05 scheduler.py

@@ -51,6 +51,29 @@ def list2map(league, l):
         m[tuple(new_k)] = new_v
     return m
 
+def binary_search(A,x,lo=0, hi=None):
+    # Binary Search that returns the index of x if x is in the array A
+    # otherwise, returns the index of minimum element that is larger than x
+    # Returns None if there is no element greater than or equal to x
+    # lo index
+    # hi index
+    if hi == None:
+        hi = len(A)-1
+    if hi - lo <= 0:
+        for i in range(lo,len(A)):
+            if A[i] >= x:
+                return i
+        return None
+    mid = (hi-lo)/2 + lo
+    mid_val = A[mid]
+    if x == mid_val:
+        return mid
+    elif x < mid_val:
+        return binary_search(A,x,lo,mid-1)
+    else:
+        # x > mid_val
+        return binary_search(A,x,mid+1, hi)
+
 class Scheduler(object):
     
     def __init__(self, year, matchups_json = False, venues_json = False):
@@ -103,13 +126,29 @@ class Scheduler(object):
                 window = 10
                 dates_domains = {}
                 dates_selected = {}
+                #print len(self.data.game_indices)
                 multiplier = len(self.data.game_indices)/initialState.TOTAL_GAMES
+                last_date = max(self.data.game_indices)
                 for state in venues:
                     t1, t2, g_n = state
                     window_min = max(0, multiplier*g_n - window)
-                    window_max = min(max(self.data.game_indices), multiplier*g_n + window)
+                    window_max = min(last_date, multiplier*g_n + window)
                     ht = t1 if venues[state] else t2
-                    dates_domains[state] = [d for d in self.data.game_indices if (d >= window_min and d <= window_max and d in ht.home_dates)]
+                    first_game_in_window_idx = binary_search(ht.home_dates,window_min)
+                    dates_domain = []
+                    #no home games after window_min
+                    # this shouldn't happen
+                    if first_game_in_window_idx is None:
+                        dates_selected[state] = dates_domain
+                    else:
+                        #add all home dates in window to dates_domain
+                        for d in xrange(first_game_in_window_idx, len(ht.home_dates)):
+                            if d <= window_max:
+                                dates_domain.append(d)
+                            else:
+                                break
+                    #dates_domains[state] = [d for d in ht.home_dates if (d >= window_min and d <= window_max)]
+                    dates_domains[state] = dates_domain
[…]

### 2015-12-05 scheduler_with_binary_search.py

@@ -1,222 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Sat Nov 21 12:24:56 2015
-
-CSP code running DFS on variables
-
-@author: charlesliu
-"""
-
-import domain as dom
-import datagen as dg
-import json
-import os
-
-def write_output(domains, filename):
-    try:
-        os.remove(filename)
-    except OSError:
-        pass
-    with open(filename + '.json', 'w') as fp:
-        json.dump(map2list(domains), fp)
-
-def read_output(league, filename):
-    with open(filename + '.json', 'r') as fp:
-        domains = list2map(league, json.load(fp))
-    return domains
-    
-def map2list(m):
-    l = []
-    for k,v in m.iteritems():
-        new_k = []
-        new_v = v.name if type(v) is dg.org.Team else v
-        for t in k:
-            if type(t) is dg.org.Team:
-                new_k.append(t.name)
-            else:
-                new_k.append(t)
-        l.append({'key':new_k, 'value': new_v})
-    return l
-    
-def list2map(league, l):
-    m = {}
-    for kv in l:
-        new_k = []
-        new_v = league.get_team(kv['value']) if type(kv['value']) is unicode else kv['value']
-        for t in kv['key']:
-            if type(t) is unicode:
-                new_k.append(league.get_team(t))
-            else:
-                new_k.append(t)
-        m[tuple(new_k)] = new_v
-    return m
-
-def binary_search(A,x,lo=0, hi=None):
-    # Binary Search that returns the index of x if x is in the array A
-    # otherwise, returns the index of minimum element that is larger than x
-    # Returns None if there is no element greater than or equal to x
-    # lo index
-    # hi index
-    if hi == None:
-        hi = len(A)-1
-    if hi - lo <= 0:
-        for i in range(lo,len(A)):
-            if A[i] >= x:
-                return i
-        return None
-    mid = (hi-lo)/2 + lo
-    mid_val = A[mid]
-    if x == mid_val:
-        return mid
-    elif x < mid_val:
-        return binary_search(A,x,lo,mid-1)
-    else:
-        # x > mid_val
-        return binary_search(A,x,mid+1, hi)
-
-class Scheduler(object):
-    
-    def __init__(self, year, matchups_json = False, venues_json = False):
-        self.data = dg.DataGen(year)
-        self.matchups_json = matchups_json
-        self.venues_json = venues_json
-
-    def create_schedule(self):
[…]

### 2015-12-06 constraint.py
venue changes

@@ -54,11 +54,9 @@ def valid_date(m, sk):
             domains[state][:] = []
     
     return True
-        
-#check that the total number of home games is TOTAL_GAMES/2
-def valid_venue(m, sk):
-    selected = m.states[m.selected]
-    domains = m.states[m.domains]
+    
+def home_away_num_game_dicts(sk, selected):
+    t1, t2, gn = sk
     home_games = {}
     away_games = {}
     num_games = {}
@@ -66,13 +64,10 @@ def valid_venue(m, sk):
     # affected by the selected game.
     # Note, since we ordered when generating team pairs, t1 < t2. So if t1, t2 are both in a game,
     # the order is (t1,t2), not (t2,t1)
-    t1, t2, gn = sk
     for state in selected:
         tx, ty, gnxy = state
         if (t1 in [tx,ty] or t2 in [tx,ty]) and (selected[state] is not None):
             home = selected[state]
-            
-            
             if t1 == tx:
                 if t2 == ty:
                     add(num_games, (t1, t2))
@@ -86,7 +81,15 @@ def valid_venue(m, sk):
                 add(home_games if home else away_games, t2)
             #t2 == ty
             else:
-                add(away_games if home else home_games, t2)   
+                add(away_games if home else home_games, t2)
+    return (home_games, away_games, num_games)
+        
+#check that the total number of home games is TOTAL_GAMES/2
+def valid_venue(m, sk):
+    selected = m.states[m.selected]
+    domains = m.states[m.domains]
+    t1, t2, gn = sk
+    home_games, away_games, num_games = home_away_num_game_dicts(sk, selected)
       
     # scheduled the required number of games between the two teams
     # clear domain

### 2015-12-06 domain.py

@@ -151,61 +151,109 @@ class Venues(TGBase):
         return min_state
     
     def min_key(self):
-        teams_selected = {}
+        '''
+            Goal is to first assign the games for teams with opponents they have an
+            even number of games with. Then once they are all assigned we try to fit
+            the remaining games, there's a little more flexibility there as it's not strict
+            split
+        '''
+        teams_selected_even = {}
+        teams_selected_odd = {}
+        #calculate number of games against even-gamed opponents
+        total_even_games = 4*self.DIVISIONAL_GAMES + 6*self.CONF_OPPONENTS_GAMES + 15*self.INTERCONF_GAMES
+        total_odd_games = self.TOTAL_GAMES - total_even_games
+        begin_search = True
         for state in self.states[self.selected]:
             if self.states[self.selected][state] is not None:
+                begin_search = False
                 t1, t2, g_n = state
+                total_games = total_even_games if constraint.total_games(self, t1, t2)%2==0 else total_odd_games
+                teams_selected = teams_selected_even if constraint.total_games(self, t1, t2)%2==0 else teams_selected_odd
                 constraint.add(teams_selected, t1)
                 constraint.add(teams_selected, t2)
-                if teams_selected[t1] == self.TOTAL_GAMES:
+                if teams_selected[t1] == total_games:
                     teams_selected.pop(t1, None)
-                if teams_selected[t2] == self.TOTAL_GAMES:
+                if teams_selected[t2] == total_games:
                     teams_selected.pop(t2, None)
         
-        #teams_selected stores how many teams have selected their home/away for games
+        #teams_selected_* stores how many teams have selected their home/away for games
         #removing those that have had all their games scheduled with home/away
         #if it's empty then no team has had a selected home/away yet
-        if len(teams_selected) == 0:
+        if begin_search:
             for k in self.states[self.domains]:
-                if len(self.states[self.domains][k]) > 0:
+                if len(self.states[self.domains][k]) > 0 and constraint.total_games(self, k[0], k[1])%2==0:
                     return (k, self.min_key_helper(k))
                     
         #choose the team with the most selected home/away venues, i.e. least domain size remaining
-        min_k = sorted(teams_selected.keys(), key=lambda x: -teams_selected[x])[0]
-        best_k = None
-        best_len = float("-inf")
+        min_k = None
+        print len(teams_selected_even)
+        if len(teams_selected_even) > 0:
+            min_k = sorted(teams_selected_even.keys(), key=lambda x: -teams_selected[x])[0]
[…]

### 2015-12-06 domain.py
fix typos

@@ -183,14 +183,17 @@ class Venues(TGBase):
             for k in self.states[self.domains]:
                 if len(self.states[self.domains][k]) > 0 and constraint.total_games(self, k[0], k[1])%2==0:
                     return (k, self.min_key_helper(k))
+        elif len(teams_selected_even) == 0 and len(teams_selected_odd) == 0:
+            for k in self.states[self.domains]:
+                if len(self.states[self.domains][k]) > 0 and constraint.total_games(self, k[0], k[1])%2!=0:
+                    return (k, self.min_key_helper(k))
                     
         #choose the team with the most selected home/away venues, i.e. least domain size remaining
         min_k = None
-        print len(teams_selected_even)
         if len(teams_selected_even) > 0:
-            min_k = sorted(teams_selected_even.keys(), key=lambda x: -teams_selected[x])[0]
+            min_k = sorted(teams_selected_even.keys(), key=lambda x: -teams_selected_even[x])[0]
         elif len(teams_selected_odd) > 0:
-            min_k = sorted(teams_selected_odd.keys(), key=lambda x: -teams_selected[x])[0]
+            min_k = sorted(teams_selected_odd.keys(), key=lambda x: -teams_selected_odd[x])[0]
         else:
             return (None, None)
         best_k_even = None

### 2015-12-07 constraint.py
updates

@@ -67,7 +67,7 @@ def home_away_num_game_dicts(sk, selected):
     for state in selected:
         tx, ty, gnxy = state
         if (t1 in [tx,ty] or t2 in [tx,ty]) and (selected[state] is not None):
-            home = selected[state]
+            date, home = selected[state]
             if t1 == tx:
                 if t2 == ty:
                     add(num_games, (t1, t2))
@@ -193,4 +193,5 @@ def game_counts(m, team):
         else:
             return None
     return game_counts
+
     

### 2015-12-07 domain.py

@@ -12,7 +12,7 @@ import constraint
 import random
 
 class TGBase(object):    
-    def __init__(self, states = None, league = None, all_dates = None):
+    def __init__(self, states = None, league = None, all_dates = None, master_dates = None):
         self.DIVISIONAL_GAMES = 4
         self.CONF_OPPONENTS_GAMES = 4
         self.REMAINING_CONF_GAMES = 3
@@ -21,6 +21,7 @@ class TGBase(object):
         self.states = {} if states is None else states
         self.domains = "domains"
         self.selected = "selected"
+        self.master_dates = "master_dates"
         if len(self.states) == 0:
             domains = {}
             selected = {}
@@ -36,6 +37,9 @@ class TGBase(object):
                     selected[(team, i)] = None
             self.states[self.domains] = domains
             self.states[self.selected] = selected
+            
+        if master_dates is not None:
+            self.selected[self.master_dates] = master_dates
             
 
     def min_key(self):
@@ -93,11 +97,14 @@ class TGBase(object):
     def copy(self):
         domains = {}
         selected = {}
+        master_dates = {}
         for k in self.states[self.domains]:
             domains[k] = self.states[self.domains][k][:]
         for k in self.states[self.selected]:
-            selected[k] = self.states[self.selected][k]
-        return {self.domains:domains, self.selected:selected}
+            selected[k] = self.states[self.selected][k][:] if self.states[self.selected][k] is not None else None
+        for k in self.states[self.master_dates]:
+            master_dates[k] = self.states[self.master_dates][k][:]
+        return {self.domains:domains, self.selected:selected, self.master_dates:master_dates}
         
 class Matchups(TGBase):
     def successors(self):
@@ -147,7 +154,7 @@ class Venues(TGBase):
             t1, t2, gn = state
             if self.states[self.selected][state] is None and t1 is tx and t2 is ty and gn < min_gn:
                     min_gn = gn
-                    min_state = state      
+                    min_state = state
         return min_state
     
     def min_key(self):
@@ -239,24 +246,56 @@ class Venues(TGBase):
                 on inconsistent states
             '''
             if t1_home == total or t2_away == total:
-                return [False] if False in self.states[self.domains][dk] else []
+                return self.domain_dates(False, sk)
             elif t1_away == total or t2_home == total:
-                return [True] if True in self.states[self.domains][dk] else []
[…]

### 2015-12-07 organization.py

@@ -33,8 +33,8 @@ class Team(OrgEq):
         skip = 5
         for i in range(len(indices))[::skip]:
             slice_len = len(indices[i:i+skip])
-            num_pick = 2 if slice_len == skip else 1
-            dates = np.random.choice(indices[i:i+5], num_pick, replace=False)
+            num_pick = 4 if slice_len == skip else 1
+            dates = np.random.choice(indices[i:i+skip], num_pick, replace=False)
             dates.sort()
             self.home_dates += list(dates)
         #these are the conference opponents that you'll play 4 games against (6 of them)

### 2015-12-07 scheduler.py

@@ -73,13 +73,22 @@ def binary_search(A,x,lo=0, hi=None):
     else:
         # x > mid_val
         return binary_search(A,x,mid+1, hi)
+        
+    
+def date_ranges(multiplier, team, game_num):
+    window = 10
+    dates = []
+    for i in range(multiplier*game_num-window, multiplier*game_num+window+1):
+        if i in team.home_dates:
+            dates.append(i)
+    return dates
 
 class Scheduler(object):
     
-    def __init__(self, year, matchups_json = False, venues_json = False):
+    def __init__(self, year, matchups_json = False, venues_dates_json = False):
         self.data = dg.DataGen(year)
         self.matchups_json = matchups_json
-        self.venues_json = venues_json
+        self.venues_dates_json = venues_dates_json
 
     def create_schedule(self):
         print "{}: Starting".format(dg.datetime.datetime.today())
@@ -98,12 +107,16 @@ class Scheduler(object):
             #(team1, team2, game_num) -> [True/False]
             #where True indicates it's played at team1's venue
             #also team1 < team2 alphabetically
-            venues = None
-            if self.venues_json:
+            venues_dates = None
+            if self.venues_dates_json:
                 venues = read_output(self.data.league, "Venues")
+                dates = read_output(self.data.league, "Dates")
+                for state in venues:
+                    venues_dates[state] = (dates[state], venues[state])
             else:
                 venues_domains = {}
                 venues_selected = {}
+                master_dates = {}
                 for state in matchups:
                     t, g_n = state
                     o = matchups[state]
@@ -111,61 +124,33 @@ class Scheduler(object):
                     dk = (t, o) if t.name < o.name else (o, t)
                     if sk not in venues_selected:
                         venues_selected[sk] = None
+                        dates_true = date_ranges(2, dk[0], g_n)
+                        dates_false = date_ranges(2, dk[1], g_n)
+                        master_dates[(sk,True)] = dates_true
+                        master_dates[(sk,False)] = dates_false
                         venues_domains[dk] = [True, False] * ((dom.constraint.total_games(initialState, t, o) + 1)/2)
-                venueState = dom.Venues({initialState.domains: venues_domains, initialState.selected: venues_selected})
-                venues, venues_statesExplored = self.DFS(venueState)
-                print "{}: Venues, {} states explored, {}".format(dg.datetime.datetime.today(), venues_statesExplored, venues is not None)
-            
-            if venues is not None:
[…]

### 2015-12-07 domain.py
reorder dates

@@ -245,10 +245,10 @@ class Venues(TGBase):
                 odd number of games against can tilt the number of home/away games
                 on inconsistent states
             '''
-            if t1_home == total or t2_away == total:
-                return self.domain_dates(False, sk)
-            elif t1_away == total or t2_home == total:
-                return self.domain_dates(True, sk)
+            if t1_home == total or t2_away == total and False in self.states[self.domains][dk]:
+                return self.domain_dates(False, sk, True)
+            elif t1_away == total or t2_home == total and True in self.states[self.domains][dk]:
+                return self.domain_dates(True, sk, True)
             '''
                 We want to try the T/F that there's more of in the domain first
                 Check for reverse because DFS takes states in reverse order
@@ -260,19 +260,34 @@ class Venues(TGBase):
             return []
             
     def order_TF(self, dk, t_first, sk):
+        true_dates = []
+        false_dates = []
         dates = []
-        domain = []
-        if t_first in self.states[self.domains][dk]: domain.append(t_first)
-        if (not t_first) in self.states[self.domains][dk]: domain.append(not t_first)
-        for t_f in domain:
-            dates += self.domain_dates(t_f, sk)
-        dates.sort(key=lambda x: -(x[0] - sk[2]*2))
+        if t_first in self.states[self.domains][dk]: true_dates = self.domain_dates(t_first, sk)
+        if (not t_first) in self.states[self.domains][dk]: false_dates = self.domain_dates(not t_first, sk)
+        
+        true_dates.sort(key=lambda x: -abs(x[0] - (sk[2]-1)*2))
+        false_dates.sort(key=lambda x: -abs(x[0] - (sk[2]-1)*2))
+        for i in range(max(len(true_dates), len(false_dates))-1, -1, -1):
+            if i >= len(true_dates):
+                dates = [false_dates[i]] + dates
+            elif i >= len(false_dates):
+                dates = [true_dates[i]] + dates
+            else:
+                if t_first:
+                    dates = [false_dates[i]] + dates
+                    dates = [true_dates[i]] + dates
+                else:
+                    dates = [true_dates[i]] + dates
+                    dates = [false_dates[i]] + dates
         return dates
         
-    def domain_dates(self, t_f, sk):
+    def domain_dates(self, t_f, sk, sort = False):
         dates = []
         for i in range(len(self.states[self.master_dates][(sk, t_f)])-1, -1, -1):
[…]

### 2015-12-07 domain.py
fix interleaving

@@ -245,9 +245,9 @@ class Venues(TGBase):
                 odd number of games against can tilt the number of home/away games
                 on inconsistent states
             '''
-            if t1_home == total or t2_away == total and False in self.states[self.domains][dk]:
+            if (t1_home == total or t2_away == total) and False in self.states[self.domains][dk]:
                 return self.domain_dates(False, sk, True)
-            elif t1_away == total or t2_home == total and True in self.states[self.domains][dk]:
+            elif (t1_away == total or t2_home == total) and True in self.states[self.domains][dk]:
                 return self.domain_dates(True, sk, True)
             '''
                 We want to try the T/F that there's more of in the domain first
@@ -268,24 +268,26 @@ class Venues(TGBase):
         
         true_dates.sort(key=lambda x: -abs(x[0] - (sk[2]-1)*2))
         false_dates.sort(key=lambda x: -abs(x[0] - (sk[2]-1)*2))
-        for i in range(max(len(true_dates), len(false_dates))-1, -1, -1):
-            if i >= len(true_dates):
-                dates = [false_dates[i]] + dates
-            elif i >= len(false_dates):
-                dates = [true_dates[i]] + dates
+        for i in range(max(len(true_dates), len(false_dates))):
+            true_index = len(true_dates) - 1 - i
+            false_index = len(false_dates) - 1 - i
+            if true_index < 0:
+                dates = [false_dates[false_index]] + dates
+            elif false_index < 0:
+                dates = [true_dates[true_index]] + dates
             else:
                 if t_first:
-                    dates = [false_dates[i]] + dates
-                    dates = [true_dates[i]] + dates
+                    dates = [false_dates[false_index]] + dates
+                    dates = [true_dates[true_index]] + dates
                 else:
-                    dates = [true_dates[i]] + dates
-                    dates = [false_dates[i]] + dates
+                    dates = [true_dates[true_index]] + dates
+                    dates = [false_dates[false_index]] + dates
         return dates
         
     def domain_dates(self, t_f, sk, sort = False):
         dates = []
-        for i in range(len(self.states[self.master_dates][(sk, t_f)])-1, -1, -1):
-            dates.append((self.states[self.master_dates][(sk, t_f)][i], t_f))
+        for i in self.states[self.master_dates][(sk, t_f)]:
+            dates.append((i, t_f))
         if sort:
             dates.sort(key=lambda x: -abs(x[0] - (sk[2]-1)*2))
         return dates

### 2015-12-07 domain.py
add randomness to t_first

@@ -11,7 +11,7 @@ Domain: (Opposing Team, Home/Away (TRUE/FALSE), DateIndex)
 import constraint
 import random
 
-class TGBase(object):    
+class TGBase(object):
     def __init__(self, states = None, league = None, all_dates = None, master_dates = None):
         self.DIVISIONAL_GAMES = 4
         self.CONF_OPPONENTS_GAMES = 4
@@ -37,10 +37,10 @@ class TGBase(object):
                     selected[(team, i)] = None
             self.states[self.domains] = domains
             self.states[self.selected] = selected
-            
+
         if master_dates is not None:
             self.selected[self.master_dates] = master_dates
-            
+
 
     def min_key(self):
         min_len = float("inf")
@@ -53,10 +53,10 @@ class TGBase(object):
         if min_k is None:
             return (None, None)
         return (min_k, self.min_key_helper(min_k))
-        
+
     def min_key_helper(self, min_k):
         raise NotImplementedError('Base class, use a super class')
-    
+
     def successorDomains(self, domain_class, constraint_func):
         domains = []
         dk, sk = self.min_key()
@@ -69,31 +69,31 @@ class TGBase(object):
                     new_obj = self.copy_states(domain_class, dk, sk, dom_elem)
                     if constraint_func(new_obj, sk):
                         domains.append(new_obj)
-                    added_elems.add(dom_elem) 
+                    added_elems.add(dom_elem)
             return domains
         return None
-    
+
     def ordered_domain(self, dk, sk):
         return self.states[self.domains][dk]
-        
+
     def copy_states(self, domain_class, dk, sk, dom_elem):
         new_states = self.copy()
         new_states[self.domains][dk].remove(dom_elem)
         new_states[self.selected][sk] = dom_elem
-        new_obj = domain_class(new_states)    
+        new_obj = domain_class(new_states)
         return new_obj
-        
+
     def successors(self):
         raise NotImplementedError('Base class, use a super class')
-        
+
     def complete(self):
         for k in self.states[self.selected]:
             if self.states[self.selected][k] is None:
                 return False
         return True
-     
+
     #replace deepcopy because it's too slow
-    #only need to copy new arrays, not objects within    
+    #only need to copy new arrays, not objects within
     def copy(self):
         domains = {}
         selected = {}
@@ -105,7 +105,7 @@ class TGBase(object):
         for k in self.states[self.master_dates]:
[…]

### 2015-12-07 scheduler.py
updates

@@ -11,6 +11,7 @@ import domain as dom
 import datagen as dg
 import json
 import os
+import constraint
 
 def write_output(domains, filename):
     try:
@@ -75,10 +76,10 @@ def binary_search(A,x,lo=0, hi=None):
         return binary_search(A,x,mid+1, hi)
         
     
-def date_ranges(multiplier, team, game_num):
+def date_ranges(multiplier, team, game_num, game_idx):
     window = 10
     dates = []
-    for i in range(multiplier*game_num-window, multiplier*game_num+window+1):
+    for i in range(max(0, multiplier*game_num-window), min(multiplier*game_num+window+1, game_idx)):
         if i in team.home_dates:
             dates.append(i)
     return dates
@@ -124,8 +125,8 @@ class Scheduler(object):
                     dk = (t, o) if t.name < o.name else (o, t)
                     if sk not in venues_selected:
                         venues_selected[sk] = None
-                        dates_true = date_ranges(2, dk[0], g_n)
-                        dates_false = date_ranges(2, dk[1], g_n)
+                        dates_true = date_ranges(2, dk[0], g_n, max(self.data.game_indices))
+                        dates_false = date_ranges(2, dk[1], g_n, max(self.data.game_indices))
                         master_dates[(sk,True)] = dates_true
                         master_dates[(sk,False)] = dates_false
                         venues_domains[dk] = [True, False] * ((dom.constraint.total_games(initialState, t, o) + 1)/2)
@@ -148,7 +149,26 @@ class Scheduler(object):
                         sched_map[(t2, g_n)] = (t1, not home, date)
                         venues[state] = home
                         dates[state] = date
-                        
+                    
+                    #sanity check that everything is right
+                    for team in self.data.league.teams():
+                        date = -20
+                        opponent_venue = {}
+                        for i in range(1, 83):
+                            opponent, t_f, d = sched_map[(team, i)]
+                            if date >= d:
+                                print "invalid date. game {} has date {}, game {} has date {}".format(i-1, date, i, d)
+                            date = d
+                            key = (opponent, t_f)
+                            constraint.add(opponent_venue, key)
+                        for opponent in self.data.league.teams():
+                            if opponent is not team:
+                                if opponent_venue[(opponent, True)] < (constraint.total_games(initialState, team, opponent))/2:
+                                    print "not enough home games"
+                                if opponent_venue[(opponent, False)] < (constraint.total_games(initialState, team, opponent))/2:
+                                    print "not enough away games"
[…]

### 2015-12-07 scheduler.py
added randomness to home_dates

@@ -24,7 +24,7 @@ def read_output(league, filename):
     with open(filename + '.json', 'r') as fp:
         domains = list2map(league, json.load(fp))
     return domains
-    
+
 def map2list(m):
     l = []
     for k,v in m.iteritems():
@@ -37,7 +37,7 @@ def map2list(m):
                 new_k.append(t)
         l.append({'key':new_k, 'value': new_v})
     return l
-    
+
 def list2map(league, l):
     m = {}
     for kv in l:
@@ -73,18 +73,24 @@ def binary_search(A,x,lo=0, hi=None):
     else:
         # x > mid_val
         return binary_search(A,x,mid+1, hi)
-        
-    
-def date_ranges(multiplier, team, game_num):
+
+
+def date_ranges(multiplier, team, game_num,last_game_idx):
     window = 10
+    #test_dates
+    #for i in range(max(0,multiplier*game_num-window), min(multiplier*game_num+window+1, last_game_idx+1)):
+    #home_dates
     dates = []
     for i in range(multiplier*game_num-window, multiplier*game_num+window+1):
+    #for i in range(max(0,multiplier*game_num-window), min(multiplier*game_num+window+1, last_game_idx+1)):
         if i in team.home_dates:
+
             dates.append(i)
+    print "team: {}, game_num: {}, dates: {}".format(team.name, game_num, dates)
     return dates
 
 class Scheduler(object):
-    
+
     def __init__(self, year, matchups_json = False, venues_dates_json = False):
         self.data = dg.DataGen(year)
         self.matchups_json = matchups_json
@@ -99,9 +105,9 @@ class Scheduler(object):
         else:
             matchups, matchups_statesExplored = self.DFS(initialState)
             print "{}: Matchups, {} states explored, {}".format(dg.datetime.datetime.today(), matchups_statesExplored, matchups is not None)
-        
+
         if matchups is not None:
-            write_output(matchups, "Matchups")        
+            write_output(matchups, "Matchups")
             #now we have our matchups which is a dict of (team, game_num) -> [opponent]
             #we next want to assign home/away so we create a map of the form:
             #(team1, team2, game_num) -> [True/False]
@@ -124,15 +130,15 @@ class Scheduler(object):
                     dk = (t, o) if t.name < o.name else (o, t)
                     if sk not in venues_selected:
                         venues_selected[sk] = None
-                        dates_true = date_ranges(2, dk[0], g_n)
[…]

### 2015-12-07 scheduler.py
reinitialize dates

@@ -77,7 +77,7 @@ def binary_search(A,x,lo=0, hi=None):
 
 
 def date_ranges(multiplier, team, game_num):
-
+    dates = []
     window = 10
     for i in range(multiplier*game_num-window, multiplier*game_num+window+1):
         if i in team.home_dates:

### 2015-12-07 organization.py
add command line output sched

@@ -33,7 +33,7 @@ class Team(OrgEq):
         skip = 5
         for i in range(len(indices))[::skip]:
             slice_len = len(indices[i:i+skip])
-            num_pick = np.random.choice([3,4,5]) if slice_len == skip else 1
+            num_pick = np.random.choice([3,4,5]) if slice_len == skip else slice_len
             dates = np.random.choice(indices[i:i+skip], num_pick, replace=False)
             dates.sort()
             self.home_dates += list(dates)

### 2015-12-07 read_schedule.py

@@ -1,0 +1,30 @@
+#!/usr/bin/python
+import sys, getopt
+import scheduler
+
+def main(argv):
+    team = ''
+    num_games = 82
+    sched = scheduler.Scheduler(2015, True, True)
+    new_sched = sched.create_schedule()
+    try:
+        opts, args = getopt.getopt(argv,"ht:n:",["team=","num_game="])
+    except getopt.GetoptError:
+        print 'read_schedule.py -t <team1,team2,...> -n <num games>'
+        sys.exit(2)
+    for opt, arg in opts:
+        if opt == '-h':
+            print 'read_schedule.py -t <team1,team2,...> -n <num games>'
+            print 'The list of possible teams are:'
+            print sorted(sched.data.league.teams(), key=lambda x: x.name)
+            print 'num games between 1 and 82'
+            print 'Output will be of the form Date,Home Team,Away Team'
+            sys.exit()
+        elif opt in ("-t", "--team"):
+            team = arg.replace(" ", "").split(",")
+        elif opt in ("-n", "--num_games"):
+            num_games = int(arg)
+    print sched.str_schedule(new_sched, num_games, team)
+
+if __name__ == "__main__":
+   main(sys.argv[1:])

### 2015-12-07 scheduler.py

@@ -30,26 +30,33 @@ def map2list(m):
     l = []
     for k,v in m.iteritems():
         new_k = []
-        new_v = v.name if type(v) is dg.org.Team else v
-        for t in k:
-            if type(t) is dg.org.Team:
-                new_k.append(t.name)
-            else:
-                new_k.append(t)
-        l.append({'key':new_k, 'value': new_v})
+        new_v = v.name if type(v) is dg.org.Team else (str(v) if type(v) is dg.datetime.date else v)
+        if type(k) is int:
+            l.append({'key':k, 'value':new_v})
+        else:
+            for t in k:
+                if type(t) is dg.org.Team:
+                    new_k.append(t.name)
+                else:
+                    new_k.append(t)
+            l.append({'key':new_k, 'value': new_v})
     return l
 
 def list2map(league, l):
     m = {}
     for kv in l:
-        new_k = []
         new_v = league.get_team(kv['value']) if type(kv['value']) is unicode else kv['value']
-        for t in kv['key']:
-            if type(t) is unicode:
-                new_k.append(league.get_team(t))
-            else:
-                new_k.append(t)
-        m[tuple(new_k)] = new_v
+        if new_v is None: new_v = kv['value']
+        if type(kv['key']) is list:
+            new_k = []
+            for t in kv['key']:
+                if type(t) is unicode:
+                    new_k.append(league.get_team(t))
+                else:
+                    new_k.append(t)
+            m[tuple(new_k)] = new_v
+        else:
+            m[kv['key']] = new_v
     return m
 
 def binary_search(A,x,lo=0, hi=None):
@@ -84,22 +91,23 @@ def date_ranges(multiplier, team, game_num):
             dates.append(i)
     return dates
 
-class Scheduler(object):
 
+debug = False
+class Scheduler(object):
     def __init__(self, year, matchups_json = False, venues_dates_json = False):
         self.data = dg.DataGen(year)
         self.matchups_json = matchups_json
         self.venues_dates_json = venues_dates_json
 
     def create_schedule(self):
-        print "{}: Starting".format(dg.datetime.datetime.today())
+        if debug: print "{}: Starting".format(dg.datetime.datetime.today())
         matchups = None
         initialState = dom.Matchups(None, self.data.league, self.data.game_indices)
         if self.matchups_json:
             matchups = read_output(self.data.league, "Matchups")
         else:
             matchups, matchups_statesExplored = self.DFS(initialState)
[…]

### 2015-12-07 scheduler.py
add row of Date,Home,Away

@@ -214,7 +214,7 @@ class Scheduler(object):
 
     #methods for printing schedules
     def str_schedule(self, schedule, games = 82, teams = []):
-        output = ""
+        output = "Date,Home,Away\n"
         if type(teams) is str:
             teams = [teams]
         for t in (teams if len(teams) > 0 else self.data.league.teams()):

### 2015-12-07 scheduler.py
revert main in scheduler

@@ -230,7 +230,7 @@ class Scheduler(object):
 
 if __name__ == '__main__':
     #The 2 True's correspond to reading in Matches.json and Venues.json so you don't have to recalculate
-    sched = Scheduler(2015,True,True)
+    sched = Scheduler(2015,True,False)
     today = dg.datetime.datetime.today()
     new_sched = sched.create_schedule()
     fin = dg.datetime.datetime.today()

### 2015-12-07 constraint.py
clean up old dates stuff

@@ -28,32 +28,6 @@ def total_games(m, team, opponent):
         return m.REMAINING_CONF_GAMES
     else:
         return m.INTERCONF_GAMES
-        
-def valid_date(m, sk):
-    domains = m.states[m.domains]
-    selected = m.states[m.selected]
-    t1, t2, g_n = sk
-    dateindex = selected[sk]
-    #constraints are:
-    #any game before g_n can't have a date >= g_n
-    #any game after g_n can't have a date <= g_n
-    
-    for state in domains:
-        x1, x2, g_x = state
-        #involves one of the two teams playing but not sk
-        if state is not sk and (t1 in [x1, x2] or t2 in [x1, x2]):
-            if g_x < g_n:
-                #remove all dates >= dateindex
-                domains[state][:] = [x for x in domains[state] if x < dateindex]
-            elif g_x > g_n:
-                #remove all dates <= dateindex
-                domains[state][:] = [x for x in domains[state] if x > dateindex]
-            if len(domains[state]) == 0 and selected[state] is None:
-                return False
-        elif state is sk:
-            domains[state][:] = []
-    
-    return True
     
 def home_away_num_game_dicts(sk, selected):
     t1, t2, gn = sk

### 2015-12-07 domain.py

@@ -314,10 +314,5 @@ class Venues(TGBase):
 
         new_obj = domain_class(new_states)
         return new_obj
-
-class Dates(TGBase):
-    def successors(self):
-        return self.successorDomains(Dates, constraint.valid_date)
-    def min_key_helper(self, min_k):
-        return min_k
+        
 

### 2015-12-07 read_schedule.py

@@ -6,6 +6,7 @@ def main(argv):
     team = ''
     num_games = 82
     sched = scheduler.Scheduler(2015, True, True)
+    sched.set_debug(False)
     new_sched = sched.create_schedule()
     try:
         opts, args = getopt.getopt(argv,"ht:n:",["team=","num_game="])

### 2015-12-07 scheduler.py

@@ -92,22 +92,25 @@ def date_ranges(multiplier, team, game_num):
     return dates
 
 
-debug = False
 class Scheduler(object):
     def __init__(self, year, matchups_json = False, venues_dates_json = False):
         self.data = dg.DataGen(year)
         self.matchups_json = matchups_json
         self.venues_dates_json = venues_dates_json
+        self.debug = True
+        
+    def set_debug(self, t):
+        self.debug = t
 
     def create_schedule(self):
-        if debug: print "{}: Starting".format(dg.datetime.datetime.today())
+        if self.debug: print "{}: Starting".format(dg.datetime.datetime.today())
         matchups = None
         initialState = dom.Matchups(None, self.data.league, self.data.game_indices)
         if self.matchups_json:
             matchups = read_output(self.data.league, "Matchups")
         else:
             matchups, matchups_statesExplored = self.DFS(initialState)
-            if debug: print "{}: Matchups, {} states explored, {}".format(dg.datetime.datetime.today(), matchups_statesExplored, matchups is not None)
+            if self.debug: print "{}: Matchups, {} states explored, {}".format(dg.datetime.datetime.today(), matchups_statesExplored, matchups is not None)
 
         if matchups is not None:
             write_output(matchups, "Matchups")
@@ -142,7 +145,7 @@ class Scheduler(object):
                         venues_domains[dk] = [True, False] * ((dom.constraint.total_games(initialState, t, o) + 1)/2)
                 venueState = dom.Venues({initialState.domains: venues_domains, initialState.selected: venues_selected, initialState.master_dates: master_dates})
                 venues_dates, venues_statesExplored = self.DFS(venueState)
-                if debug: print "{}: Venues, {} states explored, {}".format(dg.datetime.datetime.today(), venues_statesExplored, venues_dates is not None)
+                if self.debug: print "{}: Venues, {} states explored, {}".format(dg.datetime.datetime.today(), venues_statesExplored, venues_dates is not None)
 
             if venues_dates is not None:
                 venues = {}

### 2015-12-08 domain.py
Wrote A-star implementation to optimize m games in n nights constraint.  However, the heuristic needs improvement and currently, it moves at a BFS pace

@@ -12,7 +12,7 @@ import constraint
 import random
 
 class TGBase(object):
-    def __init__(self, states = None, league = None, all_dates = None, master_dates = None):
+    def __init__(self, states = None, league = None, all_dates = None, last_date = None):
         self.DIVISIONAL_GAMES = 4
         self.CONF_OPPONENTS_GAMES = 4
         self.REMAINING_CONF_GAMES = 3
@@ -22,6 +22,12 @@ class TGBase(object):
         self.domains = "domains"
         self.selected = "selected"
         self.master_dates = "master_dates"
+        self.taken_dates = 'taken_dates'
+        self.num_date_blocks = 'num_date_blocks'
+        self.current_cost = 'current_cost'
+        self.states[self.current_cost] = 0
+        if last_date is not None:
+            self.last_date = last_date
         if len(self.states) == 0:
             domains = {}
             selected = {}
@@ -37,9 +43,6 @@ class TGBase(object):
                     selected[(team, i)] = None
             self.states[self.domains] = domains
             self.states[self.selected] = selected
-
-        if master_dates is not None:
-            self.selected[self.master_dates] = master_dates
 
 
     def min_key(self):
@@ -66,12 +69,16 @@ class TGBase(object):
             added_elems = set()
             for dom_elem in self.ordered_domain(dk, sk):
                 if dom_elem not in added_elems:
+
                     new_obj = self.copy_states(domain_class, dk, sk, dom_elem)
                     if constraint_func(new_obj, sk):
                         domains.append(new_obj)
                     added_elems.add(dom_elem)
             return domains
         return None
+    def cost(self, dk, sk, dom_elem):
+        # Returns the cost of assigning dom_elem to sk
+        return 0
 
     def ordered_domain(self, dk, sk):
         return self.states[self.domains][dk]
@@ -98,13 +105,19 @@ class TGBase(object):
         domains = {}
         selected = {}
         master_dates = {}
+        taken_dates = {}
+        num_date_blocks = {}
         for k in self.states[self.domains]:
             domains[k] = self.states[self.domains][k][:]
         for k in self.states[self.selected]:
             selected[k] = self.states[self.selected][k][:] if self.states[self.selected][k] is not None else None
         for k in self.states[self.master_dates]:
             master_dates[k] = self.states[self.master_dates][k][:]
-        return {self.domains:domains, self.selected:selected, self.master_dates:master_dates}
[…]

### 2015-12-08 scheduler.py

@@ -12,6 +12,7 @@ import datagen as dg
 import json
 import os
 import constraint
+import heapq as Q
 
 def write_output(domains, filename):
     try:
@@ -98,7 +99,7 @@ class Scheduler(object):
         self.matchups_json = matchups_json
         self.venues_dates_json = venues_dates_json
         self.debug = True
-        
+
     def set_debug(self, t):
         self.debug = t
 
@@ -128,6 +129,12 @@ class Scheduler(object):
                 for state in venues:
                     venues_dates[state] = (dates[state], venues[state])
             else:
+                #dictionary: keys are teams, values are lists of dates that have been assigned
+                taken_dates = {}
+                num_date_blocks = {}
+                for team in self.data.league.teams():
+                    taken_dates[team] = []
+                    num_date_blocks[team] = 1
                 venues_domains = {}
                 venues_selected = {}
                 master_dates = {}
@@ -143,7 +150,9 @@ class Scheduler(object):
                         master_dates[(sk,True)] = dates_true
                         master_dates[(sk,False)] = dates_false
                         venues_domains[dk] = [True, False] * ((dom.constraint.total_games(initialState, t, o) + 1)/2)
-                venueState = dom.Venues({initialState.domains: venues_domains, initialState.selected: venues_selected, initialState.master_dates: master_dates})
+                venueState = dom.Venues({initialState.domains: venues_domains, initialState.selected: venues_selected,\
+                                         initialState.master_dates: master_dates, initialState.taken_dates: taken_dates, initialState.num_date_blocks: num_date_blocks}, None, None, self.data.game_indices[-1])
+                #venues_dates, venues_statesExplored = self.A_STAR(venueState)
                 venues_dates, venues_statesExplored = self.DFS(venueState)
                 if self.debug: print "{}: Venues, {} states explored, {}".format(dg.datetime.datetime.today(), venues_statesExplored, venues_dates is not None)
 
@@ -194,10 +203,6 @@ class Scheduler(object):
 
         while frontier:
             state = frontier.pop()
-            current_selected = []
-            for (k,v) in state.states[state.selected].items():
-                if v is not None:
-                    current_selected.append((k,v))
             statesExplored += 1
             if statesExplored % 1000 == 0:
                 num_selected = 0
@@ -213,6 +218,99 @@ class Scheduler(object):
                 if successors is not None:
                     frontier.extend(successors)
         return (None, statesExplored)
+
[…]

### 2015-12-08 domain.py
Changed heuristic

@@ -25,7 +25,8 @@ class TGBase(object):
         self.taken_dates = 'taken_dates'
         self.num_date_blocks = 'num_date_blocks'
         self.current_cost = 'current_cost'
-        self.states[self.current_cost] = 0
+        if self.current_cost not in self.states:
+            self.states[self.current_cost] = 1230
         if last_date is not None:
             self.last_date = last_date
         if len(self.states) == 0:
@@ -327,9 +328,10 @@ class Venues(TGBase):
                 if g_nx > g_n:
                     new_states[self.master_dates][(selected_state, True)][:] = [d for d in new_states[self.master_dates][(selected_state, True)] if d > date]
                     new_states[self.master_dates][(selected_state, False)][:] = [d for d in new_states[self.master_dates][(selected_state, False)] if d > date]
-        added_cost = self.cost(dk,sk,dom_elem)
+
         #new_states[self.current_cost] = self.states[self.current_cost] + added_cost
-        new_states[self.current_cost] = added_cost
+        new_states[self.current_cost] = self.cost(dk, sk, dom_elem)
+
         #print new_states[self.current_cost]
         new_obj = domain_class(new_states)
         # update cost
@@ -343,6 +345,7 @@ class Venues(TGBase):
         t, o, g_n = sk
         date, t_f = dom_elem
         # we assign determine cost based on sum of (maximum number of games (m) in n consecutive nights for each team)
+        m = 4
         n = 5
         # number of assignments that have already been made
         num_selected = 0
@@ -369,9 +372,13 @@ class Venues(TGBase):
                         t_num_games_in_n_nights[datex] += 1
                     elif tx is o or ox is o:
                         o_num_games_in_n_nights[datex] += 1
-        added_cost = max([max(t_num_games_in_n_nights.values()),max(o_num_games_in_n_nights.values())])
-        #print "added_cost: {}".format(added_cost)
-        return added_cost
+        max_games_in_n_nights = max([max(t_num_games_in_n_nights.values()),max(o_num_games_in_n_nights.values())])
+        total_games = len(self.states[self.selected])
+        current_cost = total_games - num_selected
+        if max_games_in_n_nights > m:
+            return current_cost + float('inf')
+        else:
+            return current_cost
 
     def update_date_blocks(self, new_states,team, date):
         team_taken_dates = new_states[self.taken_dates][team]

### 2015-12-08 scheduler.py

@@ -152,7 +152,7 @@ class Scheduler(object):
                         venues_domains[dk] = [True, False] * ((dom.constraint.total_games(initialState, t, o) + 1)/2)
                 venueState = dom.Venues({initialState.domains: venues_domains, initialState.selected: venues_selected,\
                                          initialState.master_dates: master_dates, initialState.taken_dates: taken_dates, initialState.num_date_blocks: num_date_blocks}, None, None, self.data.game_indices[-1])
-                #venues_dates, venues_statesExplored = self.A_STAR(venueState)
+                venues_dates, venues_statesExplored = self.A_STAR(venueState)
                 venues_dates, venues_statesExplored = self.DFS(venueState)
                 if self.debug: print "{}: Venues, {} states explored, {}".format(dg.datetime.datetime.today(), venues_statesExplored, venues_dates is not None)
 
@@ -238,7 +238,7 @@ class Scheduler(object):
         key = tuple(selected_list)
 
         # Initialize
-        state_cost[key] = initialState.states[initialState.current_cost]
+        state_cost[key] = 82
         priority = state_cost[key] + self.heuristic(initialState)
         Q.heappush(frontier, (priority, initialState))
 
@@ -248,17 +248,17 @@ class Scheduler(object):
             #print "frontier[0]: {}".format(frontier[0][0])
             state_priority, state = Q.heappop(frontier)
 
-            print "priority: {}".format(state_priority)
-            print "cost to state: {}".format(state.states[state.current_cost])
-            print "#" * 30
-            current_selected = []
-            selected_dates = []
-            for s in state.states[state.selected]:
-                    if state.states[state.selected][s] is not None:
-                        current_selected.append((s,state.states[state.selected][s]))
-                        selected_dates.append(state.states[state.selected][s][0])
-            selected_dates.sort()
-            print "Number assigned: {}".format(len(selected_dates))
+            #print "priority: {}".format(state_priority)
+            #print "cost to state: {}".format(state.states[state.current_cost])
+            #print "#" * 30
+            #current_selected = []
+            #selected_dates = []
+            #for s in state.states[state.selected]:
+            #        if state.states[state.selected][s] is not None:
+            #            current_selected.append((s,state.states[state.selected][s]))
+            #            selected_dates.append(state.states[state.selected][s][0])
+            #selected_dates.sort()
+            #print "Number assigned: {}".format(len(selected_dates))
[…]

### 2015-12-08 proposal.tex

@@
 
-	\section{Resources}
-	\begin{itemize}
-		\item \url{http://www.public.asu.edu/~huanliu/AI04S/project1.htm}
-		\item \url{http://www.cc.gatech.edu/~bdilkina/papers/NFLsched1.pdf}
-		\item \url{http://www.sloansportsconference.com/wp-content/uploads/2015/02/SSAC15-RP-Finalist-Alleviating-competitive-imbalance-in-NFL-schedules.pdf}
-	\end{itemize}
+	\section{Integer Linear Programming}
+	Integer linear programming is a mathematical optimization where all variables are restricted to integers.
+The objective function and the constraints are linear.
+In the NBA scheduling problem, the objective function is not defined and the solver has to solve the constraints are formulated in the problem.
+We solved this problem in 2 steps.
+The first was to attempt getting a schedule roundwise without considering the dates.
+Next we included dates and were able to create a player friendly schedule where there no more than 3 games in 5 days.
+
+	\subsection{Methodology for the roundwise schedule}
+	We create a variable for every team against every other team, for every round.
+The first part of the variable is the home team followed by the away team and the round number.
+We introduce the following constraints to the ILP
+	\begin{enumerate}
+		\item All teams play 82 games , 41 home and away
+		\item All constraints for the division, in conference and out of conference teams
+		\item No team can play more than one game in a round
+		\item Team can only play either home or away
+	\end{enumerate}
+	The above ILP was solved using the PuLP as well as the GUROBI solver.
+The PuLP solver took 40 minutes to converge to solution whereas the GUROBI solver took 6 seconds on an average.
+The schedule was verified using tests to check if all constraints were satisfied.
+
+	\subsection{Generating the player friendly schedule}
+	Using the dates generated by datagen, we formulated the ILP for a player friendly schedule.
+The variables are created for every date instead of every round.
+To reduce the number of variables we generate the variables for every possible home and away dates only.
+This reduces the number of variables by about 70\%.
+The time taken to create the constraints and to solve the ILP are greatly reduced.
+We use the following constraints.
+	\begin{enumerate}
+		\item All teams play 82 games , 41 home and away
+		\item All constraints for the division, in conference and out of conference teams
+		\item No team can play more than one game on a single date
+		\item Team can only play either home or away
+		\item A team can only play 3 or lesser matches in 5 days
+		\item A team does not play 2 away games without a day's break
+		\item A team does not play an away game right after a home game
+	\end{enumerate}
+	The above ILP was solved using the PuLP as well as the GUROBI solver.
+The PuLP solver took 1 hour to converge to solution whereas the GUROBI solver took 29 seconds on an average.
+In some cases the 4 games in 5 days was unavoidable due to mismatch in dates.
+But our solution has minimized it to a great extent and can ensure player vitality.
 

### 2015-12-08 domain.py
Fixed A-star.  It now completes in ~2 min

@@ -297,6 +297,7 @@ class Venues(TGBase):
             else:
                 dates = [false_dates[false_index]] + dates
                 dates = [true_dates[true_index]] + dates
+
         return dates
 
     def domain_dates(self, t_f, sk, sort = False):
@@ -332,7 +333,6 @@ class Venues(TGBase):
         #new_states[self.current_cost] = self.states[self.current_cost] + added_cost
         new_states[self.current_cost] = self.cost(dk, sk, dom_elem)
 
-        #print new_states[self.current_cost]
         new_obj = domain_class(new_states)
         # update cost
 
@@ -374,8 +374,8 @@ class Venues(TGBase):
                         o_num_games_in_n_nights[datex] += 1
         max_games_in_n_nights = max([max(t_num_games_in_n_nights.values()),max(o_num_games_in_n_nights.values())])
         total_games = len(self.states[self.selected])
-        current_cost = total_games - num_selected
-        if max_games_in_n_nights > m:
+        current_cost = total_games - num_selected + abs(date - (g_n-1)*2)
+        if max_games_in_n_nights >= m:
             return current_cost + float('inf')
         else:
             return current_cost

### 2015-12-08 scheduler.py

@@ -153,7 +153,7 @@ class Scheduler(object):
                 venueState = dom.Venues({initialState.domains: venues_domains, initialState.selected: venues_selected,\
                                          initialState.master_dates: master_dates, initialState.taken_dates: taken_dates, initialState.num_date_blocks: num_date_blocks}, None, None, self.data.game_indices[-1])
                 venues_dates, venues_statesExplored = self.A_STAR(venueState)
-                venues_dates, venues_statesExplored = self.DFS(venueState)
+                #venues_dates, venues_statesExplored = self.DFS(venueState)
                 if self.debug: print "{}: Venues, {} states explored, {}".format(dg.datetime.datetime.today(), venues_statesExplored, venues_dates is not None)
 
             if venues_dates is not None:
@@ -251,13 +251,13 @@ class Scheduler(object):
             #print "priority: {}".format(state_priority)
             #print "cost to state: {}".format(state.states[state.current_cost])
             #print "#" * 30
-            #current_selected = []
-            #selected_dates = []
-            #for s in state.states[state.selected]:
-            #        if state.states[state.selected][s] is not None:
-            #            current_selected.append((s,state.states[state.selected][s]))
-            #            selected_dates.append(state.states[state.selected][s][0])
-            #selected_dates.sort()
+            current_selected = []
+            selected_dates = []
+            for s in state.states[state.selected]:
+                    if state.states[state.selected][s] is not None:
+                        current_selected.append((s,state.states[state.selected][s]))
+                        selected_dates.append(state.states[state.selected][s][0])
+            selected_dates.sort()
             #print "Number assigned: {}".format(len(selected_dates))
             #print "assigned: {}".format(current_selected)
             #print "dates: {}".format(selected_dates)
@@ -291,6 +291,7 @@ class Scheduler(object):
                         # add successor state to priority queue if cost < previous cost to that state
                         if key not in state_cost or cost < state_cost[key]:
                             state_cost[key] = cost
+
                             priority = cost + self.heuristic(successor)
                             #print "new priority: {}".format(priority)
                             Q.heappush(frontier, (priority, successor))

### 2015-12-08 domain.py
A-star completed

@@ -12,7 +12,7 @@ import constraint
 import random
 
 class TGBase(object):
-    def __init__(self, states = None, league = None, all_dates = None, last_date = None):
+    def __init__(self, states = None, league = None, all_dates = None):
         self.DIVISIONAL_GAMES = 4
         self.CONF_OPPONENTS_GAMES = 4
         self.REMAINING_CONF_GAMES = 3
@@ -22,13 +22,9 @@ class TGBase(object):
         self.domains = "domains"
         self.selected = "selected"
         self.master_dates = "master_dates"
-        self.taken_dates = 'taken_dates'
-        self.num_date_blocks = 'num_date_blocks'
         self.current_cost = 'current_cost'
         if self.current_cost not in self.states:
             self.states[self.current_cost] = 1230
-        if last_date is not None:
-            self.last_date = last_date
         if len(self.states) == 0:
             domains = {}
             selected = {}
@@ -70,7 +66,6 @@ class TGBase(object):
             added_elems = set()
             for dom_elem in self.ordered_domain(dk, sk):
                 if dom_elem not in added_elems:
-
                     new_obj = self.copy_states(domain_class, dk, sk, dom_elem)
                     if constraint_func(new_obj, sk):
                         domains.append(new_obj)
@@ -114,11 +109,7 @@ class TGBase(object):
             selected[k] = self.states[self.selected][k][:] if self.states[self.selected][k] is not None else None
         for k in self.states[self.master_dates]:
             master_dates[k] = self.states[self.master_dates][k][:]
-        for k in self.states[self.taken_dates]:
-            taken_dates[k] = self.states[self.taken_dates][k][:]
-        for k in self.states[self.num_date_blocks]:
-            num_date_blocks[k] = self.states[self.num_date_blocks][k]
-        return {self.domains:domains, self.selected:selected, self.master_dates:master_dates, self.taken_dates:taken_dates, self.num_date_blocks: num_date_blocks, self.current_cost: self.states[self.current_cost]}
+        return {self.domains:domains, self.selected:selected, self.master_dates:master_dates, self.current_cost: self.states[self.current_cost]}
 
 class Matchups(TGBase):
     def successors(self):
@@ -314,9 +305,6 @@ class Venues(TGBase):
         new_states = self.copy()
         new_states[self.domains][dk].remove(t_f)
         new_states[self.selected][sk] = dom_elem
-        for team in [t,o]:
-            new_states[self.taken_dates][team].append(date)
[…]

### 2015-12-08 scheduler.py

@@ -129,12 +129,6 @@ class Scheduler(object):
                 for state in venues:
                     venues_dates[state] = (dates[state], venues[state])
             else:
-                #dictionary: keys are teams, values are lists of dates that have been assigned
-                taken_dates = {}
-                num_date_blocks = {}
-                for team in self.data.league.teams():
-                    taken_dates[team] = []
-                    num_date_blocks[team] = 1
                 venues_domains = {}
                 venues_selected = {}
                 master_dates = {}
@@ -151,7 +145,7 @@ class Scheduler(object):
                         master_dates[(sk,False)] = dates_false
                         venues_domains[dk] = [True, False] * ((dom.constraint.total_games(initialState, t, o) + 1)/2)
                 venueState = dom.Venues({initialState.domains: venues_domains, initialState.selected: venues_selected,\
-                                         initialState.master_dates: master_dates, initialState.taken_dates: taken_dates, initialState.num_date_blocks: num_date_blocks}, None, None, self.data.game_indices[-1])
+                                         initialState.master_dates: master_dates})
                 venues_dates, venues_statesExplored = self.A_STAR(venueState)
                 #venues_dates, venues_statesExplored = self.DFS(venueState)
                 if self.debug: print "{}: Venues, {} states explored, {}".format(dg.datetime.datetime.today(), venues_statesExplored, venues_dates is not None)
@@ -234,37 +228,17 @@ class Scheduler(object):
         for assignment in initialState.states[initialState.selected].items():
             if assignment[1] is not None:
                 selected_list.append(assignment)
-        #selected_list.sort()
         key = tuple(selected_list)
 
         # Initialize
         state_cost[key] = 82
-        priority = state_cost[key] + self.heuristic(initialState)
+        priority = state_cost[key]
         Q.heappush(frontier, (priority, initialState))
 
         while len(frontier) > 0:
-            #print "n_smallest: {}".format(Q.nsmallest(5, frontier))
-            #print "*"*30
-            #print "frontier[0]: {}".format(frontier[0][0])
             state_priority, state = Q.heappop(frontier)
-
-            #print "priority: {}".format(state_priority)
-            #print "cost to state: {}".format(state.states[state.current_cost])
-            #print "#" * 30
-            current_selected = []
-            selected_dates = []
-            for s in state.states[state.selected]:
-                    if state.states[state.selected][s] is not None:
-                        current_selected.append((s,state.states[state.selected][s]))
-                        selected_dates.append(state.states[state.selected][s][0])
[…]

### 2015-12-08 scheduler.py
Renamed A-star as Dijkstra's since, heuristic is no longer used.  Added doc string for Dijkstra's

@@ -146,7 +146,7 @@ class Scheduler(object):
                         venues_domains[dk] = [True, False] * ((dom.constraint.total_games(initialState, t, o) + 1)/2)
                 venueState = dom.Venues({initialState.domains: venues_domains, initialState.selected: venues_selected,\
                                          initialState.master_dates: master_dates})
-                venues_dates, venues_statesExplored = self.A_STAR(venueState)
+                venues_dates, venues_statesExplored = self.DIJKSTRAS(venueState)
                 #venues_dates, venues_statesExplored = self.DFS(venueState)
                 if self.debug: print "{}: Venues, {} states explored, {}".format(dg.datetime.datetime.today(), venues_statesExplored, venues_dates is not None)
 
@@ -213,13 +213,46 @@ class Scheduler(object):
                     frontier.extend(successors)
         return (None, statesExplored)
 
-    def A_STAR(self, initialState):
-        # Adapted from http://www.redblobgames.com/pathfinding/a-star/introduction.html
+    def DIJKSTRAS(self, initialState):
+        '''
+        Dijkstra's algorithm is used to find a schedule that satisfies
+        the m games in n nights constraint.
+        If "Violated m in n rule" in printed before a schedule is returned,
+        then there is no valid schedule that satisfies the m games in n nights.
+        The algorithm works as follow:
+            1. The initial state (where no matchups have been assigned
+                dates or venues), is assigned a cost of 1230 (the number of
+                remaining unassigned matchups)
+            2.  Note: that a successor state is the current state with one more
+                matchup assigned a date and venue.
+                Each successor state is assigned a cost, where the cost is
+                (total number of matchups = 1230) - (number of matchups assigned)
+                + abs((date of new assignment) - ((game number of new assignment)-1)*2)
+
+
+                However if at least one team plays at least m games in n nights,
+                 10000 is added to the cost.
+
+                Note: abs((date of new assignment) - ((game number of new assignment)-1)*2)
+                    is the distance between the assigned date, and the ideal date
+                    (where the ideal date is the (game number - 1) * 2)
+
+
+
+            3. The cost is used as the priority when adding a state to the
+            frontier priority queue.
+
+            Note: This cost function prioritizes schedules that have assign dates closer
+                to the ideal date and schedules that are closer to the goal
+                (i.e. more assigned games).  In addition, it gives very low priority
+                to schedules that break the m games in n nights constraint,
[…]

### 2015-12-08 scheduler.py
fixed typo

@@ -215,7 +215,7 @@ class Scheduler(object):
 
     def DIJKSTRAS(self, initialState):
         '''
-        Dijkstra's algorithm is used to find a schedule that satisfies
+        Use Dijkstra's algorithm to find a schedule that satisfies
         the m games in n nights constraint.
         If "Violated m in n rule" in printed before a schedule is returned,
         then there is no valid schedule that satisfies the m games in n nights.

### 2015-12-08 paper.tex
final paper

\documentclass{article}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{courier}
\usepackage{multicol}
\usepackage{microtype} %required to fix courier linebreaking issues
\usepackage{hyperref} % make references into links
\usepackage{amsfonts}
\hypersetup{colorlinks=true, linkcolor=blue} %make them pretty links
\usepackage[all]{hypcap} %make them link to the right place
\usepackage{url}
\makeatletter
\g@addto@macro{\UrlBreaks}{\UrlOrds}
\makeatother
\usepackage{geometry}
\geometry{legalpaper, portrait, margin=1in}

\begin{document}
	\title{CS182: NBA Scheduler}
	\author{Abhishek Malali (abhishekmalali@g.harvard.edu)\\
			Charles Liu (cliu02@g.harvard.edu)\\
			Samuel Daulton (sdaulton@g.harvard.edu)}
	\date{\today}
	\maketitle

	\section{Introduction}

	The National Basketball Association is a 30 team league that organizes an 82 game schedule for each team over a 6-month period. It showcases some of the best athletes in the world and has generated globally recognized brands. As such, it is a multi-billion dollar industry and is constantly looking for ways to improve. Recently, one such way has been to reorganize its schedule.

	\subsection{NBA Scheduling Formula}

	Currently there are two conferences (East, West) each comprised of three divisions each of which contain five teams. The 82-game schedule is currently set up as follows:

	\begin{enumerate}
		\item 4 games against the other 4 division opponents, [4x4=16 games]
		\item 4 games against 6 (out-of-division) conference opponents, [4x6=24 games]
		\item 3 games against the remaining 4 conference teams, [3x4=12 games]
		\item 2 games against teams in the opposing conference. [2x15=30 games]
	\end{enumerate}

	And of course, each team must play 41 home and 41 away games.

	\subsection{Court Availability}

	As these are actual games with required venues, each team must provide the following:

	\begin{enumerate}
		\item At least 50 dates on which their home court will be available
		\item 4 Mondays
		\item 4 Thursdays (to help TNT plan its telecasts).
	\end{enumerate}

	\subsection{Official Breaks}

	There are no games on the following days:

	\begin{enumerate}
		\item Christmas Eve
		\item All-Star Week
		\item NCAA Men's Division I Basketball Championship Game
	\end{enumerate}

    \subsection{Additional Assumptions}

    There are no non-NBA events in the venues that may further constrain a team's schedule.

	\section{Data Generation}

	We created simple classes representing teams, divisions, and conferences mimicking the structure of the league. We also created a data generation (datagen) class that generates a master list of dates - these dates conform to the scheduling requirements mentioned above. In particular, Christmas Eve is obviously December 24, we randomly select a week for an All-Star Break in February and the NCAA Men's Division I Basketball Championship is assumed to be the first Monday of April.

	Once we have our list of dates, we map them to indices ranging from 0 to the number of dates. All our subsequent calculations are based on these numbered indices. For each team, we randomly choose 3-5 dates in every 5-day period as potential home dates. We will discuss the implications of this in our results.

	\section{Integer Linear Programming}
	Integer linear programming is a mathematical optimization where all variables are restricted to integers. The objective function and the constraints are linear. In the NBA scheduling problem, the objective function is not defined and the solver has to solve the constraints as formulated in the problem. We solved this problem in 2 steps. The first was to attempt getting a schedule roundwise without considering the dates. Next we included dates and were able to create a player friendly schedule where there no more than 3 games in 5 days.

	\subsection{Methodology for the roundwise schedule}
	We create a variable for every team against every other team, for every round. The first part of the variable is the home team followed by the away team and the round number. We introduce the following constraints to the ILP
	\begin{enumerate}
		\item All teams play 82 games, 41 home and away
		\item All constraints for the division, in conference and out of conference teams
		\item No team can play more than one game in a round
		\item Team can only play either home or away
	\end{enumerate}
	The above ILP was solved using the PuLP as well as the GUROBI solver. The PuLP solver took 40 minutes to converge to solution whereas the GUROBI solver took 6 seconds on average. The schedule was verified using tests to check if all constraints were satisfied.

	\subsection{Generating the player friendly schedule}
	Using the dates generated by datagen, we formulated the ILP for a player friendly schedule. The variables are created for every date instead of every round.  To reduce the number of variables we generate the variables for every possible home and away dates only. This reduces the number of variables by about 70\%. The time taken to create the constraints and to solve the ILP are greatly reduced. We use the following constraints.
	\begin{enumerate}
		\item All teams play 82 games, 41 home and away
		\item All constraints for the division, in conference and out of conference teams
		\item No team can play more than one game on a single date
		\item Team can only play either home or away
		\item A team can only play 3 or lesser matches in 5 days
		\item A team does not play 2 away games without a day's break
		\item A team does not play an away game right after a home game
	\end{enumerate}
	The above ILP was solved using the PuLP as well as the GUROBI solver. The PuLP solver took 1 hour to converge to solution whereas the GUROBI solver took 29 seconds on an average. In some cases the 4 games in 5 days was unavoidable due to mismatch in dates, but our solution aimed to minimize it.

	\section{Constraint Satisfaction Problem}
	The ILP solution was mainly used as a speed benchmark for us. Although our solution is written in pure Python and the ILP solver most likely uses some C or Fortran bindings, we have the advantage of using heuristics to aid finding a solution.

	We modeled our problem in two sub-CSP's: one for attaining matchups and game numbers ($MSP$), and one using those matchups/game numbers to find suitable home/away and date assignments ($HSP$). The formulation from variable to domain was as follows:

	\begin{eqnarray*}
		MSP: (Team_1, game\_num) &=>& (Team_2) \\
		HSP: (Team_1, Team_2, game\_num) &=>& (Date, True/False)
	\end{eqnarray*}

	The results from $MSP$ formed the variables for $HSP$, with the ordering of the teams assigned lexicographically. An assignment of $True$ in the domain for $HSP$ referred to $Team_1$ having the home game.

	\subsection{Matchups CSP}
	Within our code, for each team we initialize a list of opposing teams with each team repeated to the number of times these two teams play. Our $MSP$ then will form the unique list of teams from this list to try assignments.

	\begin{figure}
	\caption{Example of current state during $MSP$. The number next to the opposing team corresponds to the number of games left to play against the Knicks.
	min\_count(Knicks)=1 and sum\_count(Knicks, Celtics)=2}
	\includegraphics[scale=0.6]{matchups_game_counts}
	\end{figure}

	We define two useful functions for our CSP:
	\begin{itemize}
		\item min\_count($Team$) - minimum row count over remaining games
		\item sum\_count($Team_1, Team_2$) - sum over column for $Team_2$
	\end{itemize}

	\subsubsection{Heuristics}
	There were a number of heuristics we used to solve this CSP.
	\begin{enumerate}
		\item We applied forward checking to each successor state we looked at. This consisted of checking that the min\_count for every team was at least 1 and the sum\_count over every pair of teams was at least as large as the number of games they had left to play.
		\item We chose the next team/game to assign by finding the team with the lowest min\_count and it's associated game.
		\item After choosing $t,g$ for the team/game number to next assign, we ordered the possible opponents using the least-constrained heuristic. This involved trying out the assignment and recalculating min\_count($t$). We then ordered the possible opponents in decreasing order - the opponent that kept the min\_count from dropping the most was the least constrained assignment.
	\end{enumerate}

	\subsubsection{Results}
	The CSP solves typically in 2-3 minutes with around 1600 states explored. The heuristic that was the most valuable was the least-constrained variable heuristic, as without it the CSP would not complete.

	\subsection{Dates/Venues CSP}
	Using the results from the Matchups CSP, we formulated this problem first as a DFS again. In our code, for any pair of teams we stored a list of True/False tuples according to the number of times those two teams played. If the teams played three times, there'd be two Trues and two Falses. We also created a master list of possible dates with the keys being $(Team_1,Team_2,game\_num,True/False)$. This represented the possible dates for any matchup and home/away combination (True meaning $Team_1$ was home). For the home team, let the set of home dates generated in our initial data generation process be $D$. We defined a hyperparameter $window$, and assigned the values to this dictionary as:

	\begin{eqnarray*}
		\forall x \in D: x \geq 2*game\_num - window \land x \leq 2*game\_num + window
	\end{eqnarray*}

	The rationale here is the season is typically 160-170 days, meaning ideally each game is every other day. So, we specify the domain of possible dates for a game to be a window of dates centered at that ideal date. The heuristics used were as follows:
	\begin{enumerate}
		\item Forward checking was relatively straightforward in this CSP. We remove any True/False possibilities for any pair of teams if one team has reached the limit of home or away games for that matchup. For any date $d$ we assign for some game\_num, we remove the dates prior to $d$ for any game after game\_num, and any date after $d$ for any game before game\_num for any matchup involving one of these teams.
		\item Choosing what matchup to next assign was the most perplexing part of this problem. We initially chose the team with the most assignments up to that point - essentially trying to assign all home/away matchups for a specific team at a time. The difficulty in this CSP is finding the right assignments for matchups of teams that play 3 times. There's an added complexity of having the right combination of 1 home/2 away and 2 home/1 away matchups that keep each team's total home/away games even. What was originally happening was the DFS would assign every matchup for every team until the last few, but it was impossible to find a right assignment for them because too many teams had their 3-game matchups with a given 2-home or 2-aways. There wasn't a good backtracking method implemented once that happened, so the CSP would stagnate through all possible iterations. Given that, we decided to first assign all matchups between opponents that play an even number of times and only after all those were completed do we start on the 3-game matchups. This was the key to solving this CSP.
		\item Once we had our matchup we wanted to assign, we checked which values True or False were possible - if a team had reached their limit of home/away games then there was no ordering to be done. If both teams could potentially be the home team, we chose to first try the team with the fewer number of home games be the potential home team. Then, for any Home/Away assignment, we had our master dates list with potential dates. We ordered these according to the distance to the ``ideal date'', the 2*game\_num we had used to first assign the window of possible dates. The resulting list of possible assignments was interleaved True/False assignments with their associated dates based on distance to the ``ideal date''.
		\begin{itemize}
			\item As an example suppose for $(Team_1,Team_2,game\_num=10)$, the window of home dates for $Team_1$ are $(17,19,21,23)$ and the window of home dates for $Team_2$ are $(18,20)$. We want to first try $Team_2$ as the home team because to that point it's had less home assignments, so False would take priority to True. The resulting ordering of the domain would be: $[(20,False),(19,True),(18,False),(21,True),(23,True),(17,True)]$ - ordering of 19/21 and 17/23 could be flipped based Python's sort implementation.
		\end{itemize}
	\end{enumerate}

	There's potentially no solution to our problem given the availability of home dates. This goes back to our original dates formulation - randomly selecting between 3-5 home dates every 5 days of the season for every team. If every date were possible - i.e. each team's possible home dates were every day of the season, our CSP solves in under a minute with typically under 1000 states explored. If we set the possible home dates to be 3 out of every 5 dates, our CSP solved roughly half the time. This was the rationale for us to randomly choose between 3-5 home dates, as we generally obtain solutions and they're more interesting than a game every other day.

	\subsubsection{$M$ games in $N$ nights}
	We use Dijkstra's algorithm to find a schedule that satisfies the ``no $M$ games in $N$ nights'' soft constraint. For our problem, let $M=4$ and $N=5$
    The algorithm works as follow:
    \begin{enumerate}
    	\item The initial state (where no matchups have been assigned dates or venues), is assigned a cost of 1230 (the number of remaining unassigned matchups). Note that a successor state is the current state with one more matchup assigned a date and home/away.
    	\item Each possible successor state is assigned a cost of $Unassigned + Dist$, where $Unassigned$ refers to the number of total matchups remaining to be assigned (should decrease from 1230 by 1 every time) and $Dist$ is the distance to the ideal date for the new matchup assigned (to retain the date ordering from our heuristics)
    	\item If one team of the new matchup assigned plays at least $M$ games in $N$ nights, a penalty 10000 is added to the cost
    \end{enumerate}
    The cost is used as the priority when adding a state to the frontier priority queue. Again, how we generate the home dates for our teams is crucial to determining the runtime of finding a solution (if it exists). If we give every team a home date for every day of the season, our UCS runs exactly as our previous DFS as every game will be assigned essentially the ideal date. The UCS will explore exactly the same states, but will take a bit longer to run given the added computation of calculating the maximum number of games over any $N$-day period - roughly 2 minutes. Sometimes a season is a little under 164 days, so in that case a solution is found in about 2500 states explored. With the random 3-5 days in any 5-day period, if a solution exists without breaking the soft constraint it will find it in approximately the same time, but if no such solution exists it takes an indefinite amount of time to find a solution. We believe having a good heuristic function would help with this, but were unable to come up with an admissible one that helped.

	\section{Repository}
	We felt it unnecessary to submit the code to Canvas as there are quite a few files. Our github repository is located at \url{https://github.com/chuckyouliu/NBA_Scheduler}. Basic description of important files:

	\begin{itemize}
		\item ILP\_formulation.py - As name suggests the ILP code is located here
		\item datagen.py - Formulation of the date indices
		\item organization.py - Classes corresponding to team/division/league/conference
		\item domain.py - Classes corresponding to the CSP's with functions calculating the heuristics
		\item constraint.py - Functions associated with forward checking, used by classes in domain.py
		\item scheduler.py - Main class for solving schedule, creates variables/domains that are fed into domain.py
	\end{itemize}
	Our results of an example schedule are stored in the json files. We created a simple command line tool to read from the json files by typing for example:
	\begin{center}
		python read\_schedule.py -n 5 -t Knicks,Celtics
	\end{center}
	This will output the first 5 games of the schedule for Knicks and Celtics. Use -h to see the list of possible teams.

\end{document}

### 2015-12-09 paper.tex
fix typo
fix formatting

@@
 \usepackage{geometry}
-\geometry{legalpaper, portrait, margin=1in}
+\geometry{legalpaper, portrait, margin=1.5in}
 
@@
 We want to first try $Team_2$ as the home team because to that point it's had less home assignments, so False would take priority to True.
-The resulting ordering of the domain would be: $[(20,False),(19,True),(18,False),(21,True),(23,True),(17,True)]$ - ordering of 19/21 and 17/23 could be flipped based Python's sort implementation.
+The resulting ordering of the domain would be: (20,False), (19,True), (18,False), (21,True), (23,True), (17,True) - ordering of 19/21 and 17/23 could be flipped based Python's sort implementation.
 		\end{itemize}

### 2015-12-09 paper.tex
write schedule command line tool

@@
 
-	\section{Repository}
+	\section{Running the Repository}
 	We felt it unnecessary to submit the code to Canvas as there are quite a few files.
@@
 	\end{itemize}
+	To create a schedule, you can use a command line tool that we've written called write\_schedule. An example command would be:
+	\begin{center}
+		python write\_schedule.py -m True -d True -y 2016
+	\end{center}
+	This would create a schedule for the year 2016.
+The -d stands for debug, so debug outputs would be shown, and the -m stands for if you'd like to use an already existing Matchups.json file, so only create new home/aways and dates not new matchups.
+Setting that to False would create an entirely new schedule with new matchups.
 	Our results of an example schedule are stored in the json files. We created a simple command line tool to read from the json files by typing for example:
@@
 
+	\section{Work Distribution}
+	\begin{itemize}
+	\item Data Generation - Charles created the structure of the league as well as the initial data generation class. As the project progressed changes were made by all members.
+	\item ILP - Sam and Abhishek worked on the initial matchups problem. From there, Abhishek took the lead on solving home/away's and dates as well as moving to the faster GUROBI.
+	\item CSP - Charles created the general structure of the scheduler/domain/constraint files and the sub-CSP problems. From there the more involved work (heuristics, testing) were split as follows:
+		\begin{itemize}
+			\item Matchups heuristics - Charles
+			\item Home/Aways heuristics - Charles and Sam pair coded
+			\item Dates/UCS - primarily Sam with some input on cost functions from Charles
+		\end{itemize}
+	\end{itemize}
+
 \end{document}

### 2015-12-09 domain.py

@@ -23,8 +23,6 @@ class TGBase(object):
         self.selected = "selected"
         self.master_dates = "master_dates"
         self.current_cost = 'current_cost'
-        if self.current_cost not in self.states:
-            self.states[self.current_cost] = 1230
         if len(self.states) == 0:
             domains = {}
             selected = {}
@@ -40,6 +38,11 @@ class TGBase(object):
                     selected[(team, i)] = None
             self.states[self.domains] = domains
             self.states[self.selected] = selected
+        
+        if self.current_cost not in self.states:
+            self.states[self.current_cost] = 1230
+        if self.master_dates not in self.states:
+            self.states[self.master_dates] = {}
 
 
     def min_key(self):
@@ -101,12 +104,12 @@ class TGBase(object):
         domains = {}
         selected = {}
         master_dates = {}
-        taken_dates = {}
-        num_date_blocks = {}
         for k in self.states[self.domains]:
             domains[k] = self.states[self.domains][k][:]
+
         for k in self.states[self.selected]:
-            selected[k] = self.states[self.selected][k][:] if self.states[self.selected][k] is not None else None
+            selected[k] = self.states[self.selected][k] if self.states[self.selected][k] is not None else None
+            
         for k in self.states[self.master_dates]:
             master_dates[k] = self.states[self.master_dates][k][:]
         return {self.domains:domains, self.selected:selected, self.master_dates:master_dates, self.current_cost: self.states[self.current_cost]}

### 2015-12-09 scheduler.py

@@ -203,7 +203,7 @@ class Scheduler(object):
                 for s in state.states[state.selected]:
                     if state.states[state.selected][s] is not None:
                         num_selected += 1
-                print "{}: {}, {}, {}".format(dg.datetime.datetime.today(), statesExplored, len(frontier), num_selected)
+                print "{}: {} statesExplored, {} frontier length, {} number of assignments".format(dg.datetime.datetime.today(), statesExplored, len(frontier), num_selected)
 
             if state.complete():
                 return (state.states[state.selected], statesExplored)
@@ -279,7 +279,7 @@ class Scheduler(object):
                 for s in state.states[state.selected]:
                     if state.states[state.selected][s] is not None:
                         num_selected += 1
-                print "{}: {}, {}, {}, {}".format(dg.datetime.datetime.today(), statesExplored, len(frontier), num_selected, state_priority)
+                print "{}: {} statesExplored, {} frontier length, {} num assignments, {} state cost".format(dg.datetime.datetime.today(), statesExplored, len(frontier), num_selected, state_priority)
 
             if state.complete():
                 # all dates have been assigned

### 2015-12-09 write_schedule.py

@@ -1,0 +1,33 @@
+#!/usr/bin/python
+import sys, getopt
+import scheduler
+
+def main(argv):
+    use_matchups = False
+    use_debug = False
+    use_year = scheduler.dg.datetime.datetime.today().year
+    try:
+        opts, args = getopt.getopt(argv,"hm:y:d:",["use_matchups=","use_year=", "use_debug"])
+    except getopt.GetoptError:
+        print 'write_schedule.py -m <true/false> -d <true/false> -y <year>'
+        sys.exit(2)
+    for opt, arg in opts:
+        if opt == '-h':
+            print 'write_schedule.py -m <true/false> -d <true/false> -y <year>'
+            print '-m Use the Matchups.json file already saved. New schedule will only have new dates. Default False.'
+            print '-d Include debug info of progress of CSP. Default False.'
+            print '-y What year the schedule should be for. Default current year.'
+            sys.exit()
+        elif opt in ("-m", "--use_matchups"):
+            use_matchups = arg[0].lower() == "t"
+        elif opt in ("-d", "--use_debug"):
+            use_debug = arg[0].lower() == "t"
+        elif opt in ("-y", "--use_year"):
+            use_year = int(arg)
+    sched = scheduler.Scheduler(use_year, use_matchups)
+    sched.set_debug(use_debug)
+    sched.create_schedule()
+    print "Schedule written to json. Use read_schedule to read from it."
+
+if __name__ == "__main__":
+   main(sys.argv[1:])

### 2015-12-09 ILP_formulation.py
Adding code for date problem so it can be run directly

@@ -2,11 +2,14 @@ import pulp
 import datagen as dgen
 import organization as org
 import itertools
+import numpy as np
+import json
 
 def create_var_string(value):
-    #Input in a 3-element-tuple
+    # Input in a 3-element-tuple
     return str(value[0])+'-'+str(value[1])+'-'+str(value[2])
 
+# Get a list of teams
 team_list = []
 data = dgen.DataGen(2015)
 conferences = data.league.conferences
@@ -16,63 +19,92 @@ for key in conferences.keys():
         teams = conf.divisions[div].teams
         for team in teams.keys():
             team_list.append(teams[team])
+
+
+#Generating list of possible game dates
+possible_game_dates = data.game_indices
+
 # Set number of games in the season
 num_games = 82          
 els = [tuple(x) for x in itertools.combinations(team_list, 2)]
 # Create dictionary where keys are string representations of games
-#    key: "{team1name}_{team2name}_{gamenum}"
+#    key: "{team1name}_{team2name}_{gamedate}"
 #    value: information about home team and away team and game number
 variable_dict = {}
 for key in els:
-    for idx in range(1,num_games+1):
-        variable_dict[create_var_string((key[0].name,key[1].name,idx))] = {'home':key[0].name,'away':key[1].name,'homeDiv':key[0].division.name,\
+    for date in key[0].home_dates:
+        variable_dict[create_var_string((key[0].name,key[1].name,date))] = {'home':key[0].name,'away':key[1].name,'homeDiv':key[0].division.name,\
                                                                            'homeConf':key[0].division.conference.name,\
                                                                            'awayConf':key[1].division.conference.name,\
-                                                                           'awayDiv':key[1].division.name,'value':0,'gameNum':idx}
-        variable_dict[create_var_string((key[1].name,key[0].name,idx))] = {'home':key[1].name,'away':key[0].name,'homeDiv':key[1].division.name,\
+                                                                           'awayDiv':key[1].division.name,'value':0,'posDate':date}
+    for date in key[1].home_dates:
+        variable_dict[create_var_string((key[1].name,key[0].name,date))] = {'home':key[1].name,'away':key[0].name,'homeDiv':key[1].division.name,\
                                                                            'homeConf':key[1].division.conference.name,\
                                                                            'awayConf':key[0].division.conference.name,\
[…]

### 2015-12-09 ILP_formulation.py
Corrections

@@ -169,6 +169,11 @@ for key in key_list:
         second_team = max(home,away)
         if first_team == home:
             value = True
+        else:
+            value = False
+        match["key"] = [first_team, second_team, date]
+        match["value"] = value
+        Venues.append(match)
 
 Matchup = []
 for key in schedule_dict.keys():
@@ -179,11 +184,7 @@ for key in schedule_dict.keys():
     match["key"] = [team_1,game_date]
     match["value"] = team_2
     Matchup.append(match)
-        else:
-            value = False
-        match["key"] = [first_team, second_team, date]
-        match["value"] = value
-        Venues.append(match)
+        
 
 ##Saving the list to file
 with open('Matchups_ILP.json', 'w') as outfile:

### 2015-12-10 paper.tex
remove heuristics word
Making required changes in the paper

@@
 	We create a variable for every team against every other team, for every round.
+In this part of the problem, we assume that there are 15 games every round with every team playing in that particular round.
 The first part of the variable is the home team followed by the away team and the round number.
-We introduce the following constraints to the ILP
+
+	\begin{eqnarray*}
+		variable : (Home \ Team, Away \ Team, Round \ Number)
+	\end{eqnarray*}
+
+	We introduce the following constraints to the ILP.
 	\begin{enumerate}
@@
 To reduce the number of variables we generate the variables for every possible home and away dates only.
-This reduces the number of variables by about 70\%.
-The time taken to create the constraints and to solve the ILP are greatly reduced.
-We use the following constraints.
+For every team, we create the variables for every home date for every opposition.
+Similarly, we create the variables for every away date for a particular opponent.
+
+	\begin{eqnarray*}
+		variable : (Home \ Team, Away \ Team, Game \ Date)
+	\end{eqnarray*}
+
+	This reduces the number of variables by about 70\%. The time taken to create the constraints and to solve the ILP are greatly reduced. We use the following constraints.
 	\begin{enumerate}
@@
 		\begin{itemize}
-			\item Matchups heuristics - Charles
-			\item Home/Aways heuristics - Charles and Sam pair coded
+			\item Matchups - Charles
+			\item Home/Aways - Charles and Sam pair coded
 			\item Dates/UCS - primarily Sam with some input on cost functions from Charles

### 2015-12-10 paper.tex
updates

@@
 
-	\subsubsection{$M$ games in $N$ nights}
-	We use Dijkstra's algorithm to find a schedule that satisfies the ``no $M$ games in $N$ nights'' soft constraint. For our problem, let $M=4$ and $N=5$
-    The algorithm works as follow:
+	\subsubsection{$M$ games in $N$ nights constraint}
+	To satisfy the constraint that each team plays no more than $M$ games in $N$ nights, we assign each state a priority and use Dijkstra's algorithm, instead of DFS, to find a schedule.
+For our problem, let $M=4$ and $N=5$, meaning that each team plays no more than 3 games in 5 nights.
+When the $M$ games in $N$ nights constraint is added to the HSP, the HSP is solved using Dijkstra's algorithm alone, not DFS.
+    The HSP was solved using Dijkstra's algorithm in the following way:
     \begin{enumerate}
     	\item The initial state (where no matchups have been assigned dates or venues), is assigned a cost of 1230 (the number of remaining unassigned matchups).
-Note that a successor state is the current state with one more matchup assigned a date and home/away.
+Here, a successor state is the current state with one more matchup assigned a date and home/away.
     	\item Each possible successor state is assigned a cost of $Unassigned + Dist$, where $Unassigned$ refers to the number of total matchups remaining to be assigned (should decrease from 1230 by 1 every time) and $Dist$ is the distance to the ideal date for the new matchup assigned (to retain the date ordering from our heuristics)

### 2015-12-10 paper.tex
updates

@@
     	\item The initial state (where no matchups have been assigned dates or venues), is assigned a cost of 1230 (the number of remaining unassigned matchups).
-Here, a successor state is the current state with one more matchup assigned a date and home/away.
+Here,  successor state is the current state with one more matchup assigned a date and home/away.
     	\item Each possible successor state is assigned a cost of $Unassigned + Dist$, where $Unassigned$ refers to the number of total matchups remaining to be assigned (should decrease from 1230 by 1 every time) and $Dist$ is the distance to the ideal date for the new matchup assigned (to retain the date ordering from our heuristics)
@@
 Again, how we generate the home dates for our teams is crucial to determining the runtime of finding a solution (if it exists).
-If we give every team a home date for every day of the season, our UCS runs exactly as our previous DFS as every game will be assigned essentially the ideal date.
-The UCS will explore exactly the same states, but will take a bit longer to run given the added computation of calculating the maximum number of games over any $N$-day period - roughly 2 minutes.
+If we give every team a home date for every day of the season, Dijkstra's algorithm runs exactly as our previous DFS as every game will be assigned essentially the ideal date.
+Dijkstra's algorithm will explore exactly the same states, but will take a bit longer to run given the added computation of calculating the maximum number of games over any $N$-day period - roughly 2 minutes.
 Sometimes a season is a little under 164 days, so in that case a solution is found in about 2500 states explored.
-With the random 3-5 days in any 5-day period, if a solution exists without breaking the soft constraint it will find it in approximately the same time, but if no such solution exists it takes an indefinite amount of time to find a solution.
+With the random 3-5 days in any 5-day period, if a solution exists without breaking the $M$ games in $N$ nights soft constraint it will find it in approximately the same time, but if no such solution exists it takes an indefinite amount of time to find a solution.
 We believe having a good heuristic function would help with this, but were unable to come up with an admissible one that helped.
@@
 			\item Home/Aways - Charles and Sam pair coded
-			\item Dates/UCS - primarily Sam with some input on cost functions from Charles
+			\item Dates/$M$ in $N$ - primarily Sam with some input on cost functions from Charles
 		\end{itemize}

### 2015-12-10 paper.tex
Revised paper

@@
     \subsection{Additional Assumptions}
-
     There are no non-NBA events in the venues that may further constrain a team's schedule.
@@
 	We created simple classes representing teams, divisions, and conferences mimicking the structure of the league.
-We also created a data generation (datagen) class that generates a master list of dates - these dates conform to the scheduling requirements mentioned above.
+We also created a data generation (datagen) class that generates a master list of dates that conform to the scheduling requirements mentioned above.
 In particular, Christmas Eve is obviously December 24, we randomly select a week for an All-Star Break in February and the NCAA Men's Division I Basketball Championship is assumed to be the first Monday of April.
 
-	Once we have our list of dates, we map them to indices ranging from 0 to the number of dates.
+	Once we have our list of dates, we map them to indices ranging from 0 to the total number of dates.
 All our subsequent calculations are based on these numbered indices.
@@
 	Integer linear programming is a mathematical optimization where all variables are restricted to integers.
-The objective function and the constraints are linear.
+The objective function and the constraints are linear equations.
 In the NBA scheduling problem, the objective function is not defined and the solver has to solve the constraints as formulated in the problem.
 We solved this problem in 2 steps.
-The first was to attempt getting a schedule roundwise without considering the dates.
-Next we included dates and were able to create a player friendly schedule where there no more than 3 games in 5 days.
+The first step was picking the home team, away team, and game number for each matchup, without considering dates.
+For this step, if Team 1 and Team 2 play each other in game number $x$, it is the Team 1's game $x$ and Team 2's game $x$, as opposed to the game being, for instance, game 3 for Team 1 and game 4 for Team 2.
+Once we verified correct results for generating matchups, the second step was to modify the problem to include dates.
+After completing the second problem, we created a player friendly schedule where there no more than 3 games in 5 days.
 
-	\subsection{Methodology for the roundwise schedule}
-	We create a variable for every team against every other team, for every round.
-In this part of the problem, we assume that there are 15 games every round with every team playing in that particular round.
-The first part of the variable is the home team followed by the away team and the round number.
+	\subsection{Methodology for the assigning game numbers to matchups}
+	We create a variable all combinations of Team 1, Team 2, game number. It is important to note that every team plays one game for each game number (1 through 82).
 
 	\begin{eqnarray*}
-		variable : (Home \ Team, Away \ Team, Round \ Number)
+		variable : (Home \ Team, Away \ Team, Game Number)
 	\end{eqnarray*}
@@
 		\item All constraints for the division, in conference and out of conference teams
-		\item No team can play more than one game in a round
+		\item No team can be involved in more than one matchup per game number
 		\item Team can only play either home or away
@@
 	\subsection{Generating the player friendly schedule}
-	Using the dates generated by datagen, we formulated the ILP for a player friendly schedule.
-The variables are created for every date instead of every round.
-To reduce the number of variables we generate the variables for every possible home and away dates only.
-For every team, we create the variables for every home date for every opposition.
-Similarly, we create the variables for every away date for a particular opponent.
+	Using the possible home dates for each team generated by datagen, we formulated the ILP for a player friendly schedule.
+Note that the difference here is that we created variables for every date instead of every game number.
+To reduce the number of variables, we only created variables using the possible home dates provided by the home team.
+For each pair of teams ($Team_1, Team_2$), we include variables for the possiblity that Team 1 is home and Team 2 is away and vice versa.
 
 	\begin{eqnarray*}
-		variable : (Home \ Team, Away \ Team, Game \ Date)
+		variable : (Home \ Team, Away \ Team, Date)
 	\end{eqnarray*}
 
-	This reduces the number of variables by about 70\%. The time taken to create the constraints and to solve the ILP are greatly reduced. We use the following constraints.
+	Creating variables using only the home dates provided by the home team instead of all possible dates reduces the number of variables by about 70\%.
+The optimization greatly reduced the runtime of the solver .
+We use the following constraints.
 	\begin{enumerate}
@@
 		\item Team can only play either home or away
-		\item A team can only play 3 or lesser matches in 5 days
+		\item A team can only play no more than 3 matches in 5 days
 		\item A team does not play 2 away games without a day's break
@@
 	\subsection{Matchups CSP}
-	Within our code, for each team we initialize a list of opposing teams with each team repeated to the number of times these two teams play.
+	Within our code, for each team we initialize a list of opposing teams with each team repeated the number of times the two teams play.
 Our $MSP$ then will form the unique list of teams from this list to try assignments.
@@
 This consisted of checking that the min\_count for every team was at least 1 and the sum\_count over every pair of teams was at least as large as the number of games they had left to play.
-		\item We chose the next team/game to assign by finding the team with the lowest min\_count and it's associated game.
-		\item After choosing $t,g$ for the team/game number to next assign, we ordered the possible opponents using the least-constrained heuristic.
+		\item We chose the next team to assign by finding the team with the lowest min\_count.
+		\item After choosing $t$ as the team to assign next, we ordered the possible opponents using the least-constrained heuristic.
 This involved trying out the assignment and recalculating min\_count($t$).
 We then ordered the possible opponents in decreasing order - the opponent that kept the min\_count from dropping the most was the least constrained assignment.
+		\item After choosing the two teams, we selected the matchup between the two chosen teams with the lowest game number to be the next matchup to be assigned a date
 	\end{enumerate}
@@
 In our code, for any pair of teams we stored a list of True/False tuples according to the number of times those two teams played.
-If the teams played three times, there'd be two Trues and two Falses.
+If the teams played three times, there would be two Trues and two Falses.
 We also created a master list of possible dates with the keys being $(Team_1,Team_2,game\_num,True/False)$.
@@
 For the home team, let the set of home dates generated in our initial data generation process be $D$.
-We defined a hyperparameter $window$, and assigned the values to this dictionary as:
+We defined a hyperparameter $window$ and assigned the values to this dictionary as follows, where $g$ is the game number:
 
 	\begin{eqnarray*}
-		\forall x \in D: x \geq 2*game\_num - window \land x \leq 2*game\_num + window
+		\forall x \in D: x \geq 2*g- window \land x \leq 2*g + window
 	\end{eqnarray*}
 
-	The rationale here is the season is typically 160-170 days, meaning ideally each game is every other day.
+	The rationale here is the season is typically between 160 and 170 days, meaning ideally each game is every other day.
 So, we specify the domain of possible dates for a game to be a window of dates centered at that ideal date.
@@
 We remove any True/False possibilities for any pair of teams if one team has reached the limit of home or away games for that matchup.
-For any date $d$ we assign for some game\_num, we remove the dates prior to $d$ for any game after game\_num, and any date after $d$ for any game before game\_num for any matchup involving one of these teams.
+For any date $d$ we assign for some game number $g$, we remove the dates prior to $d$ for any game after $g$ and any date after $d$ for any game before $g$ for any matchup involving one of these teams.
 		\item Choosing what matchup to next assign was the most perplexing part of this problem.
@@
 This was the key to solving this CSP.
-		\item Once we had our matchup we wanted to assign, we checked which values True or False were possible - if a team had reached their limit of home/away games then there was no ordering to be done.
+		\item Once we selected matchup to assign, we checked which values of True or False were possible: that is, which of the two teams could be the home team.
+If a team had reached their limit of home games or away games then there is no ordering to be done because there is at most one option.
 If both teams could potentially be the home team, we chose to first try the team with the fewer number of home games be the potential home team.
 Then, for any Home/Away assignment, we had our master dates list with potential dates.
-We ordered these according to the distance to the ``ideal date'', the 2*game\_num we had used to first assign the window of possible dates.
-The resulting list of possible assignments was interleaved True/False assignments with their associated dates based on distance to the ``ideal date''.
+We ordered these according to the distance to the ``ideal date'',  2*$g$, where $g$ is the game number we had used to first assign the window of possible dates.
+The resulting list contained possible assignments: alternating True/False assignments with their associated dates ordered by distance to the ``ideal date''.
 		\begin{itemize}
-			\item As an example suppose for $(Team_1,Team_2,game\_num=10)$, the window of home dates for $Team_1$ are $(17,19,21,23)$ and the window of home dates for $Team_2$ are $(18,20)$.
+			\item As an example suppose for $(Team_1,Team_2,$g$ =10)$, the window of home dates for $Team_1$ are $(17,19,21,23)$ and the window of home dates for $Team_2$ are $(18,20)$.
 We want to first try $Team_2$ as the home team because to that point it's had less home assignments, so False would take priority to True.
@@
 	There's potentially no solution to our problem given the availability of home dates.
-This goes back to our original dates formulation - randomly selecting between 3-5 home dates every 5 days of the season for every team.
+This goes back to our original dates formulation - randomly selecting between 3 and 5 home dates every 5 days of the season for every team.
 If every date were possible - i.e.
@@
 If we set the possible home dates to be 3 out of every 5 dates, our CSP solved roughly half the time.
-This was the rationale for us to randomly choose between 3-5 home dates, as we generally obtain solutions and they're more interesting than a game every other day.
+This was the rationale for us to randomly choose 3 to 5 home dates, as we generally obtain solutions and they're more interesting than a game every other day.
 

### 2015-12-10 domain.py

@@ -38,7 +38,7 @@ class TGBase(object):
                     selected[(team, i)] = None
             self.states[self.domains] = domains
             self.states[self.selected] = selected
-        
+
         if self.current_cost not in self.states:
             self.states[self.current_cost] = 1230
         if self.master_dates not in self.states:
@@ -109,7 +109,7 @@ class TGBase(object):
 
         for k in self.states[self.selected]:
             selected[k] = self.states[self.selected][k] if self.states[self.selected][k] is not None else None
-            
+
         for k in self.states[self.master_dates]:
             master_dates[k] = self.states[self.master_dates][k][:]
         return {self.domains:domains, self.selected:selected, self.master_dates:master_dates, self.current_cost: self.states[self.current_cost]}
@@ -268,7 +268,6 @@ class Venues(TGBase):
             else:
                 t_first = T_num < F_num
             return self.order_TF(dk, t_first, sk)
-            #return self.order_TF(dk, T_num < F_num, sk)
         else:
             return []
 

### 2015-12-10 paper.tex
updates, add final summary

@@
 
-	And of course, each team must play 41 home and 41 away games.
+	Each matchup must be split evenly home/away (or as close as possible in the case of 3 game matchups), and of course each team must play 41 home and 41 away games.
 
@@
 	\end{enumerate}
+
+	We don't concern ourselves as much with this, we will just randomly generate some set of home dates. More details will follow below.
 
@@
 
-	\begin{figure}
-	\caption{Example of current state during $MSP$. The number next to the opposing team corresponds to the number of games left to play against the Knicks.
-	min\_count(Knicks)=1 and sum\_count(Knicks, Celtics)=2}
-	\includegraphics[scale=0.6]{matchups_game_counts}
-	\end{figure}
-
 	We define two useful functions for our CSP:
@@
 The heuristic that was the most valuable was the least-constrained variable heuristic, as without it the CSP would not complete.
+
+	\begin{figure}
+	\caption{Example of current state during $MSP$.
+The number next to the opposing team corresponds to the number of games left to play against the Knicks.
+As Celtics still have to play Knicks 3 times, this state is inconsistent.
+	min\_count(Knicks)=1 and sum\_count(Knicks, Celtics)=2}
+	\includegraphics[scale=0.6]{matchups_game_counts}
+	\end{figure}
 
@@
 This was the key to solving this CSP.
-		\item Once we selected matchup to assign, we checked which values of True or False were possible: that is, which of the two teams could be the home team.
+		\item Once we selected the matchup to assign, we checked which values of True or False were possible: that is, which of the two teams could be the home team.
 If a team had reached their limit of home games or away games then there is no ordering to be done because there is at most one option.
@@
 We want to first try $Team_2$ as the home team because to that point it's had less home assignments, so False would take priority to True.
-The resulting ordering of the domain would be: (20,False), (19,True), (18,False), (21,True), (23,True), (17,True) - ordering of 19/21 and 17/23 could be flipped based Python's sort implementation.
+The resulting ordering of the domain would be: (20,False), (19,True), (18,False), (21,True), (23,True), (17,True) - ordering of 19/21 and 17/23 could be flipped based on Python's sort implementation.
 		\end{itemize}
@@
 We believe having a good heuristic function would help with this, but were unable to come up with an admissible one that helped.
+
+    \section{Summary}
+    We approached solving this problem in two different ways - using an ILP solver and writing our own CSP solver.
+We wanted to see if we could beat the ILP solver - but since switching to GUROBI it would seem to be infeasible using pure Python.
+However, we could reasonably solve the problem in under 5 minutes if a solution for the dates exists.
+There's certainly many more areas where we could have explored:
+
+    \begin{itemize}
+    	\item Finding a good heuristic for the $M$ games in $N$ nights constraint.
+If we could provide this then it should be possible to avoid getting stuck iterating in a BFS-like manner once the constraint gets broken.
+    	\item Extending the model to take into account more realistic scenarios - for example the Lakers and Clippers share a home arena.
+    	\item Adding to the cost function some sense of ``primetime'' games, where teams with a better record or larger market playing in those games get a higher score.
+    	\item Minimizing long road trips.
+    \end{itemize}
+
+    Generally, all of these problems involve playing around with the cost/heuristic function.
+It was a great experience in terms of using the tools we learned and applying them to a very specific problem - and it shows how knowledge about the problems helps in forming heuristics.
 
